---
title: "Combined Analyis"
author: "Nikhil Gupta"
date: "April 10, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup Libraries
```{r}
library(tswge)
library(tswgewrapped)
library(dplyr)

library(knitr)
library(kableExtra)
```

# Load Data
```{r}
data = read.csv("../data/economic_indicators_all_ex_3mo_china_inc_treas3mo.csv", header = TRUE)
data %>% glimpse()
```

```{r}
data = data %>% dplyr::select(-date)
data %>%  glimpse()
```

# Model Global Settings
```{r}
var_interest = 'gdp_change'
batch_size = 50
n.ahead = 2
```

# Train / Test Split
```{r}
data_train = data %>% dplyr::slice(1:(dplyr::n()-n.ahead))
data_test = data %>% dplyr::slice((dplyr::n()-n.ahead), dplyr::n())
```

# 1.0 Build and Compare Models

## 1.1 Univariate

```{r}
est.s <- est.arma.wge(data_train$gdp_change, 2, 0)
est.s$avar
```

```{r}
models = list("AR(2)" = list(phi = est.s$phi, vara = est.s$avar, res = est.s$res, sliding_ase = TRUE))
```

```{r}
mdl_compare_uni = ModelCompareUnivariate$new(data = data_train, var_interest = var_interest, mdl_list = models,
                                             n.ahead = n.ahead, batch_size = batch_size)

```

```{r}
mdl_compare_uni$evaluate_residuals()
```

```{r}
p = mdl_compare_uni$plot_batch_ases()
```

```{r}
p = mdl_compare_uni$plot_batch_forecasts()  
```

```{r}
mdl_compare_uni$plot_multiple_realizations()
```

```{r}
p = mdl_compare_uni$plot_simple_forecasts(zoom = 20)
```

## 1.2 VAR

```{r, fig.width=10, fig.height=10}
eda <- MultivariateEDA$new(data = data_train, var_interest = "gdp_change")
eda$plot_data(ncol = 3)
```

```{r}
# eda$plot_scatterplots()
```

### CCF Analysis

```{r, message=FALSE}
# plot the ccfs and get back the ccf table
ccf <- eda$plot_ccf_analysis(negative_only = TRUE)
# show the ccf table
ccf %>%
  select(-c(max_ccf_index_adjusted)) %>%
  kable() %>%
  kable_styling(bootstrap_options = "striped", full_width = F)
```


```{r}
lag.max = 10
  
models = list("AIC None" = list(select = "aic", trend_type = "none", lag.max = lag.max),
              "AIC Trend" = list(select = "aic", trend_type = "trend", lag.max = lag.max),
              "AIC Both" = list(select = "aic", trend_type = "both", lag.max = lag.max))
```

```{r}
mdl_build_var = ModelBuildMultivariateVAR$new(data = data_train, var_interest = var_interest,
                                              mdl_list = models, verbose = 0)
```

```{r}
mdl_build_var$summarize_build()
```

```{r}
mdl_build_var$get_recommendations()
```

```{r}
mdl_build_var$build_recommended_models()
```

```{r}
models = mdl_build_var$get_final_models(subset = 'r')
```

```{r}
# Setup Models to be compared with sliding ASE = TRUE
for (name in names(models)){
  models[[name]][['sliding_ase']] = TRUE
}
```

```{r}
# Initialize the ModelCompareMultivariateVAR object
mdl_compare_var = ModelCompareMultivariateVAR$new(data = data_train, var_interest = var_interest,
                                                  mdl_list = models, n.ahead = n.ahead, batch_size = batch_size, verbose = 1)
```

```{r}
p = mdl_compare_var$plot_batch_forecasts()
```

```{r}
p =mdl_compare_var$plot_boxplot_ases()
```

```{r}
mdl_compare_var$statistical_compare()
```

```{r}
p = mdl_compare_var$plot_batch_ases()
```

```{r}
mdl_compare_var$keep_models(mdl_names = c("AIC Both - R"))
```

```{r}
mdl_compare_var$plot_simple_forecasts()

```

## 1.3 NNFOR::mlp() Caret ####

```{r}
# Random Parallel
model = ModelBuildNNforCaret$new(data = data_train, var_interest = var_interest,
                                 search = 'random',
                                 grid = NA, tuneLength = 5,
                                 batch_size = batch_size, h = n.ahead,
                                 parallel = TRUE,
                                 seed = 1,
                                 verbose = 1)
```


```{r}
model$summarize_hyperparam_results()
model$summarize_best_hyperparams()
```

```{r}
model$plot_hyperparam_results()
```

```{r}
caret_model = model$get_final_models(subset = 'a')
```

```{r}
# Initialize the ModelCompareMultivariateVAR object
mdl_compare_mlp = ModelCompareNNforCaret$new(data = data_train, var_interest = var_interest,
                                             mdl_list = caret_model,
                                             verbose = 1)

```


# 2.0 Combine all models

```{r}
mdl_combine = ModelCombine$new(data = data_train, var_interest = var_interest,
                                 uni_models = mdl_compare_uni, var_models = mdl_compare_var, mlp_models = mdl_compare_mlp,
                                 verbose = 1)
```


## Statistical Comparison
```{r}
mdl_combine$plot_boxplot_ases()
```

```{r}
comparison = mdl_combine$statistical_compare()
```

## ASEs
```{r}
ases = mdl_combine$get_tabular_metrics()
ases %>% dplyr::filter(ASE > 20) %>% arrange(Batch)
```

```{r fig.width=12, fig.height=6}
mdl_combine$plot_batch_ases()
```

## Forecasts

### Batch Forecasts
```{r}
forecasts = mdl_combine$get_tabular_metrics(ases = FALSE)
forecasts
```

```{r fig.width=12, fig.height=6}
mdl_combine$plot_batch_forecasts()
```

### Simple Forecasts

```{r}
newxreg = data_test %>% dplyr::select(-!!var_interest)
mdl_combine$compute_simple_forecasts(lastn = FALSE, newxreg = newxreg)
```

```{r}
p = mdl_combine$plot_simple_forecasts(lastn = FALSE, newxreg = newxreg, zoom = 20) 
```

# Ensemble

## Create the ensemble model(s)
```{r}
mdl_combine$create_ensemble()
```
 
## Forecasts with Ensemble Models

```{r}
test_var_interest = data_test[var_interest]
print("Expected Values")
print(test_var_interest)
```

### Naive with combine = 'median'
```{r}
ensemble1 = mdl_combine$predict_ensemble(naive = TRUE, comb = 'median', newxreg = newxreg)
ensemble1
mean((ensemble1$ensemble - test_var_interest$gdp_change)^2)
```

### Naive with combine = 'mean'
```{r}
ensemble2 = mdl_combine$predict_ensemble(naive = TRUE, comb = 'mean', newxreg = newxreg)
ensemble2
mean((ensemble2$ensemble - test_var_interest$gdp_change)^2)
```

### glm ensemble
```{r}
ensemble3 = mdl_combine$predict_ensemble(naive = FALSE, newxreg = newxreg)
ensemble3
mean((ensemble3$ensemble - test_var_interest$gdp_change)^2)
```


