---
title: "Homework 9"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  github_document:
    toc: yes
    toc_depth: '6'
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
library(tswge)
library(tidyverse)
library(tswgewrapped)
```


# Problem 9.1

## Model 1

```{r}
phi = c(1.5, -1.21, 0.455)
theta = 0
n = 200
mean = 50

factor.wge(phi = phi)

xa = mean + gen.arma.wge(n = n, phi = phi, plot = FALSE, sn = 101)    
# px = plotts.sample.wge(xa)
```

```{r}
e = est.arma.wge(xa, p = length(phi))
e$avar
mean(xa)
```

### a

```{r}
models = list("Model 1" = list(phi = e$phi, theta = e$theta, vara = e$avar, res = e$res, sliding_ase = FALSE))

n.ahead = 20

mdl_compare = tswgewrapped::ModelCompareUnivariate$new(data = xa, mdl_list = models, n.ahead = n.ahead)
```

```{r}
mdl_compare$evaluate_residuals()
```

* Data is consistent with white noise

### b
```{r}
pt = plotts.true.wge(phi = e$phi, theta = e$theta)
px = plotts.sample.wge(xa)
```

### c
```{r}
p = mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
```

* Model forecasts settle down pretty quickly to mean so they may not do such a good job at long term forecasting

### d

```{r fig.height=8, message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = "realization", scales = 'free_y')
```

```{r message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = c("acf", "spectrum"), scales = 'fixed')
```

Very much similar to original realization

## Model 2
```{r}
phi = c(1.6, -1.7, 1.28, -0.72)
n = 200
mean = 200

factor.wge(phi = phi)

xb = mean + gen.arma.wge(n = n, phi = phi, plot = FALSE, sn = 101)    
```


```{r}
e = est.arma.wge(xb, p = length(phi))
e$avar
mean(xb)
```

### a

```{r}
models = list("Model 2" = list(phi = e$phi, theta = e$theta, vara = e$avar, res = e$res, sliding_ase = FALSE))
n.ahead = 20

mdl_compare = tswgewrapped::ModelCompareUnivariate$new(data = xb, mdl_list = models, n.ahead = n.ahead)
```

```{r}
mdl_compare$evaluate_residuals()
```

* Data is consistent with white noise

### b
```{r}
pt = plotts.true.wge(phi = e$phi, theta = e$theta)
px = plotts.sample.wge(xb)
```

### c
```{r}
p = mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
```

* Model forecasts do a pretty good job of forecasting last 20 values

### d

```{r fig.height=8, message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = "realization", scales = 'free_y')
```

```{r message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = c("acf", "spectrum"), scales = 'fixed')
```

Very much similar to original realization

## Model 3
```{r}
phi = c(1, -1.7, 0.8, -0.72)
theta = 0.9
n = 200
mean = -10

factor.wge(phi = phi)

xc = mean + gen.arma.wge(n = n, phi = phi, theta = theta, plot = FALSE, sn = 101)    
```


```{r}
e = est.arma.wge(xc, p = length(phi), q = length(theta))
factor.wge(e$theta)
e$avar
mean(xa)
```

### a

```{r}
models = list("Model 3" = list(phi = e$phi, theta = e$theta, vara = e$avar, res = e$res, sliding_ase = FALSE))
n.ahead = 20

mdl_compare = tswgewrapped::ModelCompareUnivariate$new(data = xc, mdl_list = models, n.ahead = n.ahead)
```

```{r}
mdl_compare$evaluate_residuals()
```

* Data is consistent with white noise

### b
```{r}
pt = plotts.true.wge(phi = e$phi, theta = e$theta)
px = plotts.sample.wge(xc)
```

### c
```{r}
p = mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
```

* Model forecasts settle down pretty quickly to mean so they may not do such a good job at long term forecasting

### d

```{r fig.height=8,  message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = "realization", scales = 'free_y')
```

```{r message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = c("acf", "spectrum"), scales = 'fixed')
```

Very much similar to original realization

# Problem 9.2
```{r}
data(wtcrude)
x = wtcrude
```

```{r}
px = plotts.sample.wge(x)
```

## wtcrude

* From last week, we picked ARMA(7,2)
```{r}
e = est.arma.wge(x, p = 7, q = 2)

factor.wge(e$theta)
e$avar
mean(x)
```

### a

```{r}
models = list("Model 1" = list(phi = e$phi, theta = e$theta, vara = e$avar, res = e$res, sliding_ase = FALSE))
n.ahead = 20

mdl_compare = tswgewrapped::ModelCompareUnivariate$new(data = x, mdl_list = models, n.ahead = n.ahead)
```

```{r}
mdl_compare$evaluate_residuals()
```

* Data is consistent with white noise

### b
```{r}
pt = plotts.true.wge(phi = e$phi)  # Solution forgot the theta term
pt = plotts.true.wge(phi = e$phi, theta = e$theta)
px = plotts.sample.wge(x)
```

### c
```{r}
p = mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
f = fore.arma.wge(x, phi = e$phi, theta = e$theta, n.ahead = 10, lastn = TRUE)
```

* Model forecasts settle down pretty quickly to mean so they may not do such a good job at long term forecasting

### d

```{r fig.height=8, message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = "realization", scales = 'free_y')
```

```{r message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = c("acf", "spectrum"), scales = 'fixed')
```

* Some realizations do a good job of capturing the wandering behavior but some do not.

## eco.mort

```{r}
data("eco.mort30")
x = eco.mort30
```

```{r}
px = plotts.sample.wge(x)
```

* Last week we modeled this as an ARMA(6,2)
```{r}
e = est.arma.wge(x, p = 6, q = 2)

factor.wge(e$theta)
e$avar
mean(x)
```

### a

```{r}
models = list("Model 1" = list(phi = e$phi, theta = e$theta, vara = e$avar, res = e$res, sliding_ase = FALSE))
n.ahead = 20

mdl_compare = tswgewrapped::ModelCompareUnivariate$new(data = x, mdl_list = models, n.ahead = n.ahead)
```

```{r}
mdl_compare$evaluate_residuals()
```


* Data is consistent with white noise

### b
```{r}
pt = plotts.true.wge(phi = e$phi)  # Solution forgot the theta term
pt = plotts.true.wge(phi = e$phi, theta = e$theta)
px = plotts.sample.wge(x)
```
* These are similar

### c
```{r}
p = mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
f = fore.arma.wge(x, phi = e$phi, theta = e$theta, n.ahead = 10, lastn = TRUE)
```

* Model forecasts settle down pretty quickly to mean so they may not do such a good job at long term forecasting

### d


```{r fig.height=8, message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = "realization", scales = 'free_y')
```

```{r message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = c("acf", "spectrum"), scales = 'fixed')
```


## hadley

```{r}
data("hadley")
x = hadley
```

```{r}
px = plotts.sample.wge(x)
```

* Last week we modeled this as an ARMA(4,2)
```{r}
e = est.arma.wge(x, p = 4, q = 2)

factor.wge(e$theta)
e$avar
mean(x)
```

### a

```{r}
models = list("Model 1" = list(phi = e$phi, theta = e$theta, vara = e$avar, res = e$res, sliding_ase = FALSE))
n.ahead = 20

mdl_compare = tswgewrapped::ModelCompareUnivariate$new(data = x, mdl_list = models, n.ahead = n.ahead)
```

```{r}
mdl_compare$evaluate_residuals()
```

* Data is consistent with white noise

### b
```{r}
pt = plotts.true.wge(phi = e$phi, theta = e$theta)
px = plotts.sample.wge(x)
```
* These are similar

### c
```{r}
p = mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
f = fore.arma.wge(x, phi = e$phi, theta = e$theta, n.ahead = 10, lastn = TRUE)
```

* Model forecasts settle down pretty quickly to mean so they may not do such a good job at long term forecasting

### d


```{r fig.height=8, message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = "realization", scales = 'free_y')
```

```{r message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = c("acf", "spectrum"), scales = 'fixed')
```

* The sample autocorrelations tend to damped much more quickly than the actual realization so maybe a non-stationary model may be appropriate here.

# Problem 9.3

```{r}
phi = c(0.8)
d = 0
theta = 0
s = 6
mean = 25
n = 200

x = gen.aruma.wge(n = n, phi = phi, theta = theta, d = d, s = s, plot = FALSE, sn = 101)
```

```{r}
px = plotts.sample.wge(x)
```

### Overfit
```{r} 
e.burg = est.ar.wge(x, p = 10, type = 'burg')
```

```{r}
factor.wge.season(6)
```

* Yes, s = 6 is present in the overfit table

### Remove s = 6 seasonality
```{r}
x.d6 = artrans.wge(x, phi.tr = c(rep(0,5),1))
```

* Resulting signal looks stationary

## Model ID
```{r}
aic5.wge(x.d6)
```

```{r}
e = est.arma.wge(x.d6, p = 1, q = 1)
factor.wge(e$theta)
e$avar
mean(x)
```

### a

```{r}
models = list("Model 1" = list(d = d, s = s, phi = e$phi, theta = e$theta, vara = e$avar, res = e$res, sliding_ase = FALSE))
n.ahead = 20

mdl_compare = tswgewrapped::ModelCompareUnivariate$new(data = x, mdl_list = models, n.ahead = n.ahead)
```

```{r}
mdl_compare$evaluate_residuals()
```

* Data is consistent with white noise

### b
```{r}
## Important ##
factors = mult.wge(fac1 = c(rep(0,5), 0.999), fac2 = e$phi)
phi = factors$model.coef
pt = plotts.true.wge(phi = phi, theta = e$theta)
px = plotts.sample.wge(x)
```
* These are similar

### c
```{r}
p = mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = FALSE)
f = fore.aruma.wge(x, d = d, s = s, phi = e$phi, theta = e$theta, n.ahead = 10, lastn = TRUE, limits = FALSE)
```

* Model forecasts do a pretty good job of forecasting the values

### d



```{r fig.height=8, message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = "realization", scales = 'free_y')
```

```{r message=FALSE, warning=FALSE}
mdl_compare$plot_multiple_realizations(n.realizations = 4, seed = 100, plot = c("acf", "spectrum"), scales = 'fixed')
```


```{r}

```



