---
title: "Homework 9"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  github_document:
    toc: yes
    toc_depth: '6'
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
library(tswge)
library(tidyverse)
library(tswgewrapped)
```


# 8.1

```{r}
data("prob8.1a")
data("prob8.1b")
data("prob8.1c")
data("prob8.1d")

```

```{r}
px = plotts.sample.wge(prob8.1a)
px = plotts.sample.wge(prob8.1b)
px = plotts.sample.wge(prob8.1c)
px = plotts.sample.wge(prob8.1d)

acf(prob8.1a)
acf(prob8.1b)
acf(prob8.1c)
acf(prob8.1d)
```

* We expect maybe 1 our of 20 ACF values to exceed the limit for white noise at maximum
* 'a' seems to be white noise
* 'b' Looks looks like AR(1) with a small negative phi value. There is some doubt that this is white noise
* 'c' Does not look like white noise. The variance is higher in the 1st hald and there seems to be an autocorelation in the second half.
* 'd' Seems consistent with white noise.

# 8.2

## a and b

### Model 1
```{r}
phi = c(1.5, -1.21, 0.455)
n = 200
mean = 50

factor.wge(phi = phi)

xa = mean + gen.arma.wge(n = n, phi = phi, plot = FALSE, sn = 101)    
px = plotts.sample.wge(xa)
```

### Model 2
```{r}
phi = c(1.6, -1.7, 1.28, -0.72)
n = 200
mean = 200

factor.wge(phi = phi)

xb = mean + gen.arma.wge(n = n, phi = phi, plot = FALSE, sn = 101)    
px = plotts.sample.wge(xb)
```

### Model 3
```{r}
phi = c(1, -1.7, 0.8, -0.72)
theta = 0.9
n = 200
mean = -10

factor.wge(phi = phi)

xc = mean + gen.arma.wge(n = n, phi = phi, theta = theta, plot = FALSE, sn = 101)    
px = plotts.sample.wge(xc)
```

## c
```{r}
aic5.wge(xa, p = 0:8, q = 0:3)
aic5.wge(xb, p = 0:8, q = 0:3)
aic5.wge(xc, p = 0:8, q = 0:3)
```

## d
### Model 1

```{r}
e.mle.a = est.arma.wge(xb, p = 3, q = 0)
mean(xa)
e.mle.a$avar
```

* Very close to original parameters

### Model 2

```{r}
e.mle.b = est.arma.wge(xb, p = 4, q = 0)
mean(xb)
e.mle.b$avar
```

* Very close to original parameters

### Model 3
```{r}
e.mle.c = est.arma.wge(xb, p = 5, q = 1)
e.mle.c$theta
mean(xc)
e.mle.c$avar

```

* Phi is very close to original parameters; theta not so much
* Extra term in phi is not so dominant.


## e

Fitted factors are fairly close to the actual ones.

# 8.3

## Setup
```{r}
data('wtcrude')
data("eco.mort30")
data("hadley")
```

## a
```{r}
px = plotts.sample.wge(wtcrude)
```

```{r}
aic5.wge(wtcrude, p = 0:8, q = 0:3)
```

```{r}
e.wtcrude = est.arma.wge(wtcrude, p = 7, q = 2)

factor.wge(e.wtcrude$theta)
mean(wtcrude)
e.wtcrude$avar
```

* Note that dominant factor is f = 0.
* The other factor with freq at 0.2758 almost gets cancelled by the MA term.



## b
```{r}
px = plotts.sample.wge(eco.mort30)
```

```{r}
aic5.wge(eco.mort30, p = 0:8, q = 0:3)
```

```{r}
e.mort = est.arma.wge(eco.mort30, p = 6, q = 2)

factor.wge(e.mort$theta)
mean(eco.mort30)
e.mort$avar
```

* MOstly wandering behavior due to f = 0
* Freq at 0.34 is almost cancelled between phi and theta
* Rest of the roots are less dominant.

## c
```{r}
px = plotts.sample.wge(hadley)
```

```{r}
aic5.wge(hadley, p = 0:8, q = 0:3)
```

```{r}
e.temp = est.arma.wge(hadley, p = 4, q = 2)

factor.wge(e.temp$theta)
mean(hadley)
e.temp$avar
```

* Wandering behavior is evident f = 0
* f = 0.5 gets cancelled between AR and MA side


# 8.5

```{r}
phi = c(0.8)
d = 0
theta = 0
s = 6
mean = 25
n = 200

x = gen.aruma.wge(n = n, phi = phi, theta = theta, d = d, s = s, plot = FALSE)
```

```{r}
px = plotts.sample.wge(x)
```

## (i)
```{r} 
e.burg = est.ar.wge(x, p = 10, type = 'burg')
```

```{r}
factor.wge.season(6)
```

* Yes, s = 6 is present in the overfit table

## (ii)
```{r}
x.d6 = artrans.wge(x, phi.tr = c(rep(0,5),1))
```

* Resulting signal looks stationary

## (iii)
```{r}
aic5.wge(x.d6)
```

```{r}
e.mle = est.ar.wge(x.d6, p = 1, type = 'mle')
e.mle$avar
mean(x)
```

* Very close to original estimates.

# 8.6

## a
```{r}
data("freight")
x = freight
```

```{r}
px = plotts.sample.wge(x)
```

* Looks to have seasonality of 12
```{r}
e.burg = est.ar.wge(x, p = 15, type = 'burg')
```

```{r}
factor.wge.season(12)
```

* s = 12 is reasonable from the overfit table.
* Also has a extra term at f = 0 indicative of trend (ARIMA with d = 1)

* Lets take out the seasonality first

```{r}
x.d12 = artrans.wge(x, phi.tr = c(rep(0,11),1))
```

* The differenced data actually shows AR properties (positive phi with wandering behavior) and not really ARIMA type of properties, so we will stop here and model as an ARMA process.

```{r}
aic5.wge(x.d12)
```

```{r}
e.mle = est.ar.wge(x.d12, p = 2, type = 'mle')
e.mle$avar
mean(x)
```

* Note that the solution actually choose to use the ARIMA term.

## b

```{r}
data("freeze")
x = freeze
```

```{r}
px = plotts.sample.wge(x, lag.max = 100)
```

* Looks to be more of a sinusoidal behavior more than seasonality.
```{r}
e.burg = est.ar.wge(x, p = 50, type = 'burg')
```

```{r}
factor.wge.season(36)
```

* Most factors can be found in the overfit tables 
* We will go with s = 36


```{r}
x.d36 = artrans.wge(x, phi.tr = c(rep(0,35),1))
```

* The differenced data actually looks like white noise.

```{r}
aic5.wge(x.d36)
```

We will go with ARMA(0,0) - White Noise

```{r}
e.mle = est.arma.wge(x.d36, p = 0)
e.mle$avar
mean(x)
```

## c
```{r}
data("lavon")
x = lavon
```

```{r}
px = plotts.sample.wge(x)
```


```{r}
e.burg = est.ar.wge(x, p = 8, type = 'burg')
```

```{r}
factor.wge.season(4)
```

* Does not seem to have a stong seasonal factor in there, so we fit a stationary model

```{r}
aic5.wge(x, p = 0:8, q = 0:3)
```

```{r}
e.mle = est.arma.wge(x, p = 6, q = 3)
factor.wge(e.mle$theta)
e.mle$avar
mean(x)
```

* Terms cancel out, so we use the next available model.

```{r}
e.mle = est.arma.wge(x, p = 8, q = 1)
factor.wge(e.mle$theta)
e.mle$avar
mean(x)
```

* This is also much closer to the spectral density of data 

* Also trying out the simper model from the 1st one (ARMA(6,3) --> AR(3,0))
```{r}
e.mle = est.arma.wge(x, p = 3, q = 0)
# factor.wge(e.mle$theta)
e.mle$avar
mean(x)
```

* This one is also close but misses out on some smaller frequencies.