---
title: "Homework 9"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  github_document:
    toc: yes
    toc_depth: '6'
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
library(tswge)
library(tidyverse)
library(tswgewrapped)
```


# Problem 7.3

## a

```{r}
data("sunspot.classic")
sunspot.classic %>%  glimpse()
x = sunspot.classic
```

```{r}
px = plotts.sample.wge(x)
```

```{r}
mean(x)
```

```{r}
e.mle.ar2 = est.ar.wge(x, p = 2, type = 'mle')
e.burg.ar2 = est.ar.wge(x, p = 2, type = 'burg')
e.yw.ar2 = est.ar.wge(x, p = 2, type = 'yw')

e.mle.ar2$avar
e.burg.ar2$avar
e.yw.ar2$avar

```

```{r}
e.mle.ar8 = est.ar.wge(x, p = 8, type = 'mle')
e.burg.ar8 = est.ar.wge(x, p = 8, type = 'burg')
e.yw.ar8 = est.ar.wge(x, p = 8, type = 'yw')

e.mle.ar8$avar
e.burg.ar8$avar
e.yw.ar8$avar
```

## b
```{r}
x = c(26.8, 27.8, 30.0, 31.6, 33.0, 34.2, 34.1, 33.2, 31.7, 31.5, 31.9, 31.5, 30.0, 29.4, 29.9)
px = plotts.sample.wge(x)
```

```{r}
mean(x)
```

```{r}
e.mle.ar1 = est.ar.wge(x, p = 1, type = 'mle')
e.mle.ar1$avar
```

## c
```{r}
e.mle.ar2 = est.ar.wge(x, p = 2, type = 'mle')
e.mle.ar2$avar
```

# 7.4

## a
```{r}
types = c('mle', 'burg', 'yw')
phi = c(1.3, -0.6)
n = 100
mean = -20
repeats = 3

factor.wge(phi = phi)

results = tribble(~actual_phi, ~est_type, ~avar, ~model_mean, ~phis) 

for (i in 1:repeats){
  p = length(phi)
  x = mean + gen.arma.wge(n = n, phi = phi, plot = FALSE)    
  
  for (type in types){
    e = est.ar.wge(x, p = p, type = type)
    results = results %>% add_row(actual_phi = paste(round(phi, 4), collapse = " "), est_type = type,
                                  avar = e$avar, model_mean = mean(x),
                                  phis = paste(round(e$phi, 4), collapse = " "))
  }
  
}

results
```

* All three are fairly equal in estimating

## b
```{r}
types = c('mle', 'burg', 'yw')
phi = c(3.1, -4.1, 2.63, -0.72)
n = 100
mean = 100
repeats = 3

factor.wge(phi = phi)

results = tribble(~actual_phi, ~est_type, ~avar, ~model_mean, ~phis) 

for (i in 1:repeats){
  p = length(phi)
  x = mean + gen.arma.wge(n = n, phi = phi, plot = FALSE)    
  
  for (type in types){
    e = est.ar.wge(x, p = p, type = type)
    results = results %>% add_row(actual_phi = paste(round(phi, 4), collapse = " "), est_type = type,
                                  avar = e$avar, model_mean = mean(x),
                                  phis = paste(round(e$phi, 4), collapse = " "))
  }
  
}

results
```

* Yule Walker estimates are not good.

# 7.5
```{r}
x = sunspot.classic
```

```{r}
px = plotts.sample.wge(x)
```

```{r}
e.mle.ar2 = est.ar.wge(x, p = 2, type = 'mle')
pt = plotts.true.wge(phi = e.mle.ar2$phi)
```

```{r}
e.mle.ar8 = est.ar.wge(x, p = 8, type = 'mle')
pt = plotts.true.wge(phi = e.mle.ar8$phi)
```

* The high frequency bumps present in the AR(8) estimates are missing in the AR(2) estimates. These bumps were seen in the original data as well.

# 7.6

## a
```{r}
types = c('mle', 'burg', 'yw')
phi = c(1.3, -0.6)
n = 100
mean = -20

factor.wge(phi = phi)

x = mean + gen.arma.wge(n = n, phi = phi, plot = FALSE)    
px = plotts.sample.wge(x)

for (type in types){
  e = est.ar.wge(x, p = p, type = type)
  pt = plotts.true.wge(phi = e$phi)
}
  
```

* All three estimates are fairly close

## b
```{r}
types = c('mle', 'burg', 'yw')
phi = c(3.1, -4.1, 2.63, -0.72)
n = 100
mean = 100

factor.wge(phi = phi)

x = mean + gen.arma.wge(n = n, phi = phi, plot = FALSE)    
px = plotts.sample.wge(x)

for (type in types){
  e = est.ar.wge(x, p = p, type = type)
  pt = plotts.true.wge(phi = e$phi)
}
  
```

* According to the spectral densities, there seem to be only a single frequency in all 3 estimaes. But the factor tabl tells a different story. MLE and Burg have 2 frequencies very close to each other, while YW only had 1 frequency at 0.09.

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

