---
title: "Multivariate Compare"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
library(tswge)
library(tswgewrapped)
library(tidyverse)
library(tseries)
library(orcutt)
library(vars)
library(RColorBrewer)
```

```{r}
data = read.csv("../Datasets/la_cmort_study.csv")
data %>% glimpse()
data %>%  skimr::skim()
```

## CCF Analysis

```{r}
lag.max = 10
```


```{r}
# By convention, the X2 variable comes first
c = ccf(data$temp, data$cmort, lag.max = lag.max)
index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
max_ccf = c$lag[,1,1][index] 
max_ccf
```

* At lag of -1, we have the maximum correlation and it seems to be negative. As temp increases, mortality decreases.

```{r}
# By convention, the X2 variable comes first
c = ccf(data$part, data$cmort, lag.max = lag.max)
index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
max_ccf = c$lag[,1,1][index] 
max_ccf
```

* Lag of 7 seems to have the maximum correlation to mortality and positive. More the particles, more the mortality.

* We will use lag.max = 10 in VARselect to give it enough room to pick the best model.

## Multivariate Compate Object
```{r}
models = list("VARS AIC No Trend" = list(type = "VAR", select = "aic", trend_type = "none", lag.max = lag.max),
              "VARS AIC Trend"    = list(type = "VAR", select = "aic", trend_type = "trend", lag.max = lag.max),
              "VARS AIC Const + Trend" = list(type = "VAR", select = "aic", trend_type = "both", lag.max = lag.max),
              "VARS BIC No Trend" = list(type = "VAR", select = "bic", trend_type = "none", lag.max = lag.max),
              "VARS BIC Trend"    = list(type = "VAR", select = "bic", trend_type = "trend", lag.max = lag.max),
              "VARS BIC Const + Trend" = list(type = "VAR", select = "bic", trend_type = "both", lag.max = lag.max)
               )
```

```{r}
n.ahead = 20
var_interest = 'cmort'
```

```{r}
all_data = data %>% dplyr::select(-Week)
```

```{r}
mdl_compare = ModelCompareMultivariate$new(data = all_data, mdl_list = models, var_interest = 'cmort',
                                           n.ahead = 20, batch_size = 52)
```

### AIC BIC Full Data
```{r}
mdl_compare$get_xIC() 
```

### Tabular Metrics
```{r}
mdl_compare$get_tabular_metrics()
mdl_compare$get_tabular_metrics(ases = FALSE) %>% 
  na.omit()
```

### Simple Forecasts
```{r fig.width=10}
mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = FALSE)
mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
mdl_compare$plot_simple_forecasts(lastn = FALSE, limits = FALSE)
mdl_compare$plot_simple_forecasts(lastn = FALSE, limits = TRUE)
```


### Batch Forecasts
```{r fig.width=10}
mdl_compare$plot_batch_forecasts(only_sliding = FALSE)
```

### Batch ASEs
```{r}
mdl_compare$plot_batch_ases(only_sliding = FALSE)
```

### ASE Histograms
```{r}
mdl_compare$plot_histogram_ases()  # Same function as univariate
```


### Statistical Compare
```{r}
mdl_compare$statistical_compare() 
```


# For pre-live

## Setup



```{r}
plot(data)
```

* There seems to be some weekly trend in the data (downward)



## Train Test Split

```{r}
tail(data)
```


```{r}
n = nrow(data)
train_data = all_data %>% dplyr::filter(row_number() <= (n - n.ahead))
test_data = all_data %>%  dplyr::filter(row_number() > (n - n.ahead))
```


## VARselect
```{r}
# VARselect will select the best K for this VAR model
# Looks at the cross correlation structure (ccf) to figure this out 
vselect = VARselect(all_data, lag.max = lag.max, type = "both", season = NULL, exogen = NULL)
vselect # Gives AIC values for various K values

k = vselect$selection[["AIC(n)"]]  # BIC = SC(n)
k
```

AIC picks a VAR(10) model

## VAR Model

```{r}
lsfit = VAR(train_data, p=k, type="both")
stats::AIC(lsfit)
a = summary(lsfit)
#a
a$varresult[var_interest]  # $coefficients[,"Pr(>|t|)"]
```

## Predictions

```{r}
preds = predict(lsfit, n.ahead=20)

results = preds$fcst[[var_interest]] %>% 
  as.tibble() %>% 
  mutate(Week = seq(n-n.ahead+1,n,1)) 

results

```

### ASE

```{r}
ASE_data = data %>% 
  dplyr::full_join(results, by = "Week") %>% 
  na.omit()

ASE_data
```

```{r}
ASE = mean((ASE_data$cmort - ASE_data$fcst)^2, na.rm = TRUE)
ASE
```


### Plots

```{r}
results = results %>% 
  dplyr::mutate(cmort = fcst) %>% 
  dplyr::mutate(Model = 'VAR Model') %>% 
  dplyr::select(Model, Week, cmort)

combined_data = data %>% 
  mutate(Model = 'Realization') %>% 
  dplyr::select(Model, Week, cmort ) %>% 
  bind_rows(results) %>% 
  mutate(Model = as.factor(Model))

combined_data
```


```{r fig.width=10}
p = ggplot2::ggplot() + 
  ggplot2::geom_line(data = combined_data, mapping = ggplot2::aes(x = Week, y = cmort, color = Model))

print(p)

plot(seq(1,n,1), data$cmort, type = "b")
points(seq(n-n.ahead+1,n,1), preds$fcst$cmort[,'fcst'], type = "b", pch = 15)
```


```{r fig.width=10}
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues")) # Change color pallet to make distinguishable. 
```




```{r}

```

