---
title: "Live Session Assignment 6"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
library(tswge)
library(dplyr)
library(tidyquant)
source("common_functions.R")
```


# For pre-live

## 2
```{r}
s = 7; factor.wge(c(rep(0,s-1), 1))   
```

## 3
### A

```{r}
# (1 - B^4)Xt = (1 + 0.3B)(1 - 0.6B + 0.8B^2)at

# (1 - B^4) -> s = 4
# There are no (1-B)^d term -> d = 0
# There is no dependence on X with lag, hence there are no AR components -> p = 0
# There are 3 roots in the characteristic equation for the MA side -> q = 3

# Hence, this is a seasonal ARUMA model with d = 0, s = 4, p = 0  and q = 3
```

### B
```{r}
# (1 + 0.3B)Xt = (1 - 0.6B + 0.8B^2)at

theta = c(0.6, -0.8)
factor.wge(phi = theta)

# There are no seasonal (1-B^s) terms -> s = 0
# There are no (1-B)^d term -> d = 0
# There is 1 root of the characteristic equation of the AR side. There are 2 roots of the characteristic equation of the MA side (and none of the factors on the AR and MA side cancel) -> -> p = 1, q = 2

# Hence, this is a ARMA(1,2) model
```

### C
```{r}
# (1 - 0.1B - 0.99B^2 + 0.013B^3 + 0.2078B^4 + 0.0888B^5 + 0.00864B^6) * (1-B)^2 * (1-B^12) * Xt = (1 - 0.6B + 0.8B^2) * at

phi = c(0.1, 0.99, -0.13, -0.2078, -0.0888, -0.00864)
factor.wge(phi = phi)

theta = c(0.6, -0.8)
factor.wge(phi = theta)

# There is a seasonal component (1-B^s) -> s = 12
# There is a (1-B)^d term -> d = 2
# There are 6 roots of the characteristic equation of the AR side. There are 2 roots of the characteristic equation of the MA side (and none of the factors on the AR and MA side cancel) -> p = 6, q = 2

# Hence, this is a Seasonal ARUMA model with p = 6, d = 2, q = 2 and s = 12
```

## 4
```{r}
txn = tidyquant::tq_get("TXN", get = "stock.prices", from = "2019-01-01")
txn %>%
  glimpse()
```


```{r}
px = plotts.sample.wge(txn$adjusted)
```

```{r}
dif1 = artrans.wge(txn$adjusted, phi.tr = 1)
px = plotts.sample.wge(dif1)
aic5.wge(dif1)
```

* 'dif1' looks to be just white noise. Hence the model best suited for this could be a simple ARIMA(0,1,0)
* The Spectral Density does indicate a few small peaks and dips, indicating that there may be some AR and MA components as well. After checking the AIC values, a fit with p = 2 and q = 0 does give the best AIC and hence we could also use ARIMA(2,1,0) and compare the results of the 2 models.


# Live Session
```{r}
# If we are trying to fit a 1-B^12 model, we fit to order 15 and then ask for phis and see if we have factors close to those that are seen in pure 1-B^12 model. 
factor.wge(c(0.6996, 0.2599, 0.0079, -0.0646, 0.1381, -0.0953, 0.0235, -0.0969, 0.1770, -0.1191, 0.1030, 0.7754, -0.4590, -0.4099, 0.0501))

factor.wge(c(rep(0,11),1))

# At least 6 of the 7 of the factors are well represented in our AR(15) model
# 1-B is not well represented but that is still OK. It can still be considered pretty close to 1-B^12
```


```{r}
x = gen.arma.wge(n = 200, phi = c(0.3, 0.5))
aic.wge(x)  # Returns the phi and theta values as well 
# aic5.wge also fits the models, but does not return the phi and theta values.
```


```{r}
factor.wge(c(0.4, 0, 0.74))
```


