---
title: "Live Session Assignment 4"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
library(tswge)
library(dplyr)
source("common_functions.R")
```

# Problem 1

```{r}
phi = c(-0.5, -0.6)
p = plotts.true.wge(phi = phi)
factor.wge(phi = phi)

```

# Problem 2

```{r}
walmart = read.csv("../Datasets/Walmart.csv")
head(walmart)
```

```{r}
stor9_item50 = walmart %>% dplyr::filter(item == 50 & store == 9)
summary(stor9_item50)
```
```{r}
p = plotts.sample.wge(stor9_item50$sales)
```


```{r}
x = stor9_item50$sales
n = length(x)
n.ahead = 30
results = tribble(~model, ~ASE) # Placeholder for results
```

## Modeling

* Check if it compares to a factor table with seasonality of 7
```{r}
factor.wge(phi = c(rep(0,6),1))
```

* Yes, it does match with factor table for seasonality of 7. 
* Let's remove it to see what is left over

```{r}
dif7 = artrans.wge(x, phi.tr = c(rep(0,6),1))
px = plotts.sample.wge(dif7)
```

* Looks like the weekly seasonality is the only component. Rest looks mainly noise (magnitude of peaks is very low). Lets check.

```{r}
aic5.wge(dif7)
```

* Looks like the peaks and dips

```{r}
fit = aic.wge(dif7)
factor.wge(fit$phi) # AR Component
factor.wge(fit$theta) # MA Component

```

* AR component models the peaks at 0.26 and 0.34 (less dominant)
* MA component models the dip at 0.28 (dominant)

## Assignment

* Now we move to the actual assignment for this week which is to see which of the ARMA models below gives the best fit.

### Model 1 Factors
```{r}
phi = c(0.967)
factor.wge(phi = phi) 
p = plotts.true.wge(phi = phi)

f = fore.arma.wge(x, phi=phi, n.ahead=n.ahead, limits=FALSE, lastn = TRUE)

# ASE
ase = mean((x[(n-n.ahead+1):n] - f$f)^2)
ase

results = results %>% add_row(model = "Model 1", ASE = ase)

# Checking whether within limits
comparison = cbind(f$f, f$ll, f$ul, x[(n-n.ahead+1):n])
comparison = data.frame(comparison)
names(comparison) = c("Forecast", "Lower Limit", "Upper Limit", "Actual")
comparison = comparison %>% 
  mutate(ll_satisfied = ifelse(Actual >= `Lower Limit`, TRUE, FALSE),
         ul_satisfied = ifelse(Actual <= `Upper Limit`, TRUE, FALSE),
         within_limits = ifelse(ll_satisfied & ul_satisfied, TRUE, FALSE))

comparison

```


### Model 2 Factors
```{r}
phi = c(1.452, -0.453, -0.294, 0.175, 0.237, -0.154)
factor.wge(phi = phi) 
p = plotts.true.wge(phi = phi)

f = fore.arma.wge(x, phi=phi, n.ahead=n.ahead, limits=FALSE, lastn = TRUE)

# ASE
ase = mean((x[(n-n.ahead+1):n] - f$f)^2)
ase

results = results %>% add_row(model = "Model 2", ASE = ase)

# Checking whether within limits
comparison = cbind(f$f, f$ll, f$ul, x[(n-n.ahead+1):n])
comparison = data.frame(comparison)
names(comparison) = c("Forecast", "Lower Limit", "Upper Limit", "Actual")
comparison = comparison %>% 
  mutate(ll_satisfied = ifelse(Actual >= `Lower Limit`, TRUE, FALSE),
         ul_satisfied = ifelse(Actual <= `Upper Limit`, TRUE, FALSE),
         within_limits = ifelse(ll_satisfied & ul_satisfied, TRUE, FALSE))

comparison

```

### Model 3 Factors
```{r}
phi = c(1.445, -0.411, -0.038, 0.17, 0.362, -0.245, -0.177, 0.213)
factor.wge(phi = phi) # Not stationary
# p = plotts.true.wge(phi = phi)
```

### Model 4 Factors
```{r}
phi = c(1.384, -0.359, -0.309, 0.063, 0.317, -0.14, -0.0587, -0.199, 0.2877)
factor.wge(phi = phi) 
p = plotts.true.wge(phi = phi)

f = fore.arma.wge(x, phi=phi, n.ahead=n.ahead, limits=FALSE, lastn = TRUE)

# ASE
ase = mean((x[(n-n.ahead+1):n] - f$f)^2)
ase

results = results %>% add_row(model = "Model 4", ASE = ase)

# Checking whether within limits
comparison = cbind(f$f, f$ll, f$ul, x[(n-n.ahead+1):n])
comparison = data.frame(comparison)
names(comparison) = c("Forecast", "Lower Limit", "Upper Limit", "Actual")
comparison = comparison %>% 
  mutate(ll_satisfied = ifelse(Actual >= `Lower Limit`, TRUE, FALSE),
         ul_satisfied = ifelse(Actual <= `Upper Limit`, TRUE, FALSE),
         within_limits = ifelse(ll_satisfied & ul_satisfied, TRUE, FALSE))

comparison

```

### Model 5 Factors
```{r}
phi = c(0.1516, 0.0769, 0.0016, -0.0095, 0.0542, 0.1117, 0.5683)
factor.wge(phi = phi) 
p = plotts.true.wge(phi = phi)

f = fore.arma.wge(x, phi=phi, n.ahead=n.ahead, limits=FALSE, lastn = TRUE)

# ASE
ase = mean((x[(n-n.ahead+1):n] - f$f)^2)
ase

results = results %>% add_row(model = "Model 5", ASE = ase)

# Checking whether within limits
comparison = cbind(f$f, f$ll, f$ul, x[(n-n.ahead+1):n])
comparison = data.frame(comparison)
names(comparison) = c("Forecast", "Lower Limit", "Upper Limit", "Actual")
comparison = comparison %>% 
  mutate(ll_satisfied = ifelse(Actual >= `Lower Limit`, TRUE, FALSE),
         ul_satisfied = ifelse(Actual <= `Upper Limit`, TRUE, FALSE),
         within_limits = ifelse(ll_satisfied & ul_satisfied, TRUE, FALSE))

comparison

```

### Model 6 Factors
```{r}
phi = c(.4295, .1769, .0358, .1454)

factor.wge(phi = phi) 
p = plotts.true.wge(phi = phi)

f = fore.arma.wge(x, phi=phi, n.ahead=n.ahead, limits=FALSE, lastn = TRUE)

# ASE
ase = mean((x[(n-n.ahead+1):n] - f$f)^2)
ase

results = results %>% add_row(model = "Model 6", ASE = ase)

# Checking whether within limits
comparison = cbind(f$f, f$ll, f$ul, x[(n-n.ahead+1):n])
comparison = data.frame(comparison)
names(comparison) = c("Forecast", "Lower Limit", "Upper Limit", "Actual")
comparison = comparison %>% 
  mutate(ll_satisfied = ifelse(Actual >= `Lower Limit`, TRUE, FALSE),
         ul_satisfied = ifelse(Actual <= `Upper Limit`, TRUE, FALSE),
         within_limits = ifelse(ll_satisfied & ul_satisfied, TRUE, FALSE))

comparison

```

### Model 7 Factors
```{r}
phi = c(0.1529, 0.0988, 0.0059, -0.0075, 0.0538, 0.1118, 0.5717, 0.0034, -0.0382 )
factor.wge(phi = phi) 
p = plotts.true.wge(phi = phi)

f = fore.arma.wge(x, phi=phi, n.ahead=n.ahead, limits=FALSE, lastn = TRUE)

# ASE
ase = mean((x[(n-n.ahead+1):n] - f$f)^2)
ase

results = results %>% add_row(model = "Model 7", ASE = ase)

# Checking whether within limits
comparison = cbind(f$f, f$ll, f$ul, x[(n-n.ahead+1):n])
comparison = data.frame(comparison)
names(comparison) = c("Forecast", "Lower Limit", "Upper Limit", "Actual")
comparison = comparison %>% 
  mutate(ll_satisfied = ifelse(Actual >= `Lower Limit`, TRUE, FALSE),
         ul_satisfied = ifelse(Actual <= `Upper Limit`, TRUE, FALSE),
         within_limits = ifelse(ll_satisfied & ul_satisfied, TRUE, FALSE))

comparison

```

```{r}
results
```

```{r}
x = gen.arma.wge(n = 200, phi = phi, sn = 101)
p = plotts.sample.wge(x)
xc = butterworth.wge(x, order = 4, type = "stop", cutoff = c(0.05, 0.11))
p = plotts.sample.wge(xc$x.filt)
```

# Breakout 4
```{r}
phi = c(0.15, -0.4)
factor.wge(phi = phi) 
p = plotts.true.wge(phi = phi)
```

```{r}
phi = c(-0.15, -0.4)
factor.wge(phi = phi) 
p = plotts.true.wge(phi = phi)
```


```{r}
phi = c(-0.9, 0.67, -0.2994)
factor.wge(phi = phi) 
# p = plotts.true.wge(phi = phi)
```