---
title: "Live Session Assignment 10"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
library(tswge)
library(tidyverse)
library(tswgewrapped)
library(orcutt)
```


# For pre-live

## Setup

```{r}
electrictiy = read.csv("../Datasets/electricity_usage_nikhil_house.csv")
head(electrictiy)
```

```{r}
sum(is.na(electrictiy$Usage))
```


* There are some missing values. We will need to impute these

```{r}
data = ts(electrictiy$Usage, frequency = 12, start = c(2009, 12))

library(imputeTS)
x = imputeTS::na_interpolation(data, option = 'spline')
x
```


```{r}
px = plotts.sample.wge(x)
```

# ARIMA Model

## Overfit table

* We see mild wandering behavior, but more than that data shows seasonality with s = 12. To verify this, we will use overfit table with s = 15

```{r}
factor.wge.season(12)
```

```{r}
e.burg = est.ar.wge(x, p = 15, type = 'burg')
```

* Just considering ARIMA, we dont see a very strong 1-B factor. It is present with a Abs Reciprocal of 0.91, so we could still consider removing it.

```{r}
x.d1 = artrans.wge(x, phi.tr = 1)
px = plotts.sample.wge(x.d1)
```

* THis shows a more sinusoidal behavior. It also looks sort of stationary. We will try to model it using aic5

## ARMA Estimation
```{r}
## Increased limit since lot of p = 4 and 5 were showing up in top 5. 
## Also used BIC for a smaller model
aic5.wge(x.d1, p = 0:8, q = 0:3, type = 'bic') 
```

```{r}
e = est.arma.wge(x.d1, p = 8, q = 1)
factor.wge(e$theta)
e$avar
```

* Although there are a couple of weak factors, the 2 main ones match the frequency peaks that show up in the Spectral Density of the differenced data. Also, the theta value points to a dip at f = 0 which again matches with what we have in the Spectral Density.
* Hence we will stick to this model.

* (1-B)(1 -0.4833B +0.1660B2 +0.2849B3 +0.2112B4 +0.0093B5 +0.1058B6 +0.1334B7 +0.3555B8) Xt = (1-0.8984B)at
* (1-B)(1-1.6869B+0.9362B^2)(1-0.8750B+0.8908B^2)(1+1.4843B+0.6649B^2)(1+0.5943B+0.6411B^2) Xt = (1-0.8984B)at

## Prediction
```{r}
f = fore.aruma.wge(x, phi = e$phi, theta = e$theta, d = 1, s = 0, n.ahead = 12, limits = FALSE)
f = fore.aruma.wge(x, phi = e$phi, theta = e$theta, d = 1, s = 0, n.ahead = 12, limits = FALSE, lastn = TRUE)
```

```{r}
r = tswgewrapped::sliding_ase_univariate(x, phi = e$phi, theta = e$theta, d = 1, s = 0, n.ahead = 12, batch_size = 60)
mean(r$ASEs)
sd(r$ASEs)
```



# Seasonal ARIMA Model

## Overfit Tables

* We see mild wandering behavior, but more than that data shows seasonality with s = 12. To verify this, we will use overfit table with s = 15

```{r}
factor.wge.season(12)
```


```{r}
e.burg = est.ar.wge(x, p = 15, type = 'burg')
```


```{r}
e.burg = est.ar.wge(x, p = 24, type = 'burg')
```

* I only see a few of the factors show up from s = 12. 
* The only factors that are appreciable are 1-1.7321B+1.0000B^2 and 1-1.0000B+1.0000B^2
* Hence only these should be used to transform the data

```{r}
factors = mult.wge(fac1 = c(1.732, -1), fac2 = c(1, -1))
factors$model.coef

factor.wge(factors$model.coef)

```

```{r}
x.tr = artrans.wge(x, phi.tr = factors$model.coef)
px = plotts.sample.wge(x.tr)
```

* This looks very much stationary. Lets model this using AIC5

```{r}
aic5.wge(x.tr, p = 0:10, q = 0:3, type = 'bic')
```

## ARMA Estimation
```{r}
e = est.arma.wge(x.tr, p = 6, q = 2)
factor.wge(e$theta)
e$avar
```

## Prediction
```{r}
f = fore.aruma.wge(x, phi = e$phi, theta = e$theta, d = 0, s = 0, lambda = factors$model.coef, n.ahead = 12, limits = FALSE)
f = fore.aruma.wge(x, phi = e$phi, theta = e$theta, d = 0, s = 0, lambda = factors$model.coef, n.ahead = 12, limits = FALSE, lastn = TRUE)
```

```{r}
r = tswgewrapped::sliding_ase_univariate(x, phi = e$phi, theta = e$theta, d = 0, s = 0, n.ahead = 12, batch_size = 60, lambda = factors$model.coef)
mean(r$ASEs)
sd(r$ASEs)
```

# Signal Plus Noise Model

```{r}
t = seq(1,length(x),1)
df = data.frame(x = x, t= t)
fit = lm(x~t, data = df)
summary(fit)
```

* Does not show a linear trend.

```{r}
cfit = cochrane.orcutt(fit) 
summary(cfit)
```

* Does not show a lineear trend in the data. Since the trend is not fouund, we will not continue the fitting process using this method.
```{r}

```


# Live Session

```{r}
# Time Series Unit 10 Live Session

# Setup
library(tswge)
data(patemp)
```

```{r}
# EDA
p = plotts.wge(patemp)
p = parzen.wge(patemp)
```

```{r}
# Signal + Noise Model
fore.patemp.spn = fore.sigplusnoise.wge(patemp, freq = .083, max.p = 4, n.ahead = 24, linear = FALSE, lastn = TRUE)
fore.patemp.spn$b
factor.wge(fore.patemp.spn$phi.z)
fore.patemp.spn$wnv

ASE.spn = mean((patemp[(length(patemp)-23): length(patemp)]-fore.patemp.spn$f)^2)
ASE.spn
```


```{r}
# ARUMA with s = 12 only
fore.patemp.B12 = fore.aruma.wge(patemp,s = 12, n.ahead = 24, lastn = TRUE)
ASE.B12 = mean((patemp[(length(patemp)-23): length(patemp)]-fore.patemp.B12$f)^2)
ASE.B12
```

```{r}
# ARUMA with s = 12 and ARMA noise
patemp_B12 = artrans.wge(patemp, phi.tr = c(rep(0,11),1))
aic5.wge(patemp_B12)
patemp.est.AR2 = est.arma.wge(patemp,p = 2)
fore.patemp.B12.AR2 = fore.aruma.wge(patemp,s = 12, phi = patemp.est.AR2$phi, n.ahead = 24, lastn = TRUE)
ASE.B12.AR2 = mean((patemp[(length(patemp)-23): length(patemp)]-fore.patemp.B12.AR2$f)^2)
ASE.B12.AR2
```

```{r}
# ARUMA with 2 Lambda terms and ARMA noise
y.tr=artrans.wge(patemp,phi.tr=c(1.732,-1))
patemp.est.lamda.AR3 = est.ar.wge(y.tr,p=3)
fore.patemp.L2.AR3 = fore.aruma.wge(patemp,lambda = c(1.732, -1), phi = patemp.est.lamda.AR3$phi, n.ahead = 24, lastn = TRUE)
ASE.L2.AR3 = mean((patemp[(length(patemp)-23): length(patemp)]-fore.patemp.L2.AR3$f)^2)
ASE.L2.AR3
```

