---
title: "Live Session Assignment 11"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  github_document:
    toc: yes
    toc_depth: 6
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r}
library(tswge)
library(tswgewrapped)
library(tidyverse)
library(tseries)
library(orcutt)
library(vars)
library(RColorBrewer)
```


# CA Mortality Read Data
```{r}
data = read.csv("../Datasets/la_cmort_study.csv")
data %>% glimpse()
data %>%  skimr::skim()
```

```{r fig.width=10}
df.m = reshape::melt(data, id.vars = "Week")
ggplot(df.m) + 
  facet_wrap(variable ~ ., ncol = 1, scales = 'free_y') + 
  geom_line(aes(x = Week, y = value, colour = variable))

```

```{r}
all_data = data %>% dplyr::select(-Week)
```

# CCF Analysis

```{r}
lag.max = 10
```

```{r}

var_interest = 'cmort'

for (name in colnames(all_data)){
  # By convention, the X2 variable comes first
  c = ccf(data[name], data[var_interest], lag.max = lag.max)
  index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
  max_ccf = c$lag[,1,1][index] 
  cat(paste("\nFor variable: ", name, " , max cross-correlation is at lag: ", max_ccf, sep = ""))
}


```

* At lag of -1, we have the maximum correlation and it seems to be negative. As temp increases, mortality decreases.
* Lag of -7 seems to have the maximum correlation to mortality and positive. More the particles, more the mortality.

* We will use lag.max = 10 in VARselect to give it enough room to pick the best model.


# Using tswgewrapped

## Multivariate Compare Object
```{r}

models = list(
  "VARS AIC Both B" = list(select = "aic", trend_type = "both", lag.max = lag.max, sliding_ase = TRUE),
  "VARS AIC Trend B" = list(select = "aic", trend_type = "trend", lag.max = lag.max, sliding_ase = TRUE),
  "VARS AIC Const B" = list(select = "aic", trend_type = "const", lag.max = lag.max, sliding_ase = TRUE),
  "VARS AIC None B" = list(select = "aic", trend_type = "none", lag.max = lag.max, sliding_ase = TRUE),
  "VARS BIC Both B" = list(select = "bic", trend_type = "both", lag.max = lag.max, sliding_ase = TRUE),
  "VARS BIC Trend B" = list(select = "bic", trend_type = "trend", lag.max = lag.max, sliding_ase = TRUE),
  "VARS BIC Const B" = list(select = "bic", trend_type = "const", lag.max = lag.max, sliding_ase = TRUE),
  "VARS BIC None B" = list(select = "bic", trend_type = "none", lag.max = lag.max, sliding_ase = TRUE)
)
```

```{r}
n.ahead = 20
batch_size = 200 
```


```{r}
mdl_compare = ModelCompareMultivariateVAR$new(data = all_data, mdl_list = models, var_interest = var_interest,
                                              n.ahead = n.ahead, batch_size = batch_size, verbose = 1)
```

### AIC BIC Full Data
```{r}
mdl_compare$get_xIC() 
```

### Tabular Metrics

```{r}
ASEs = mdl_compare$get_tabular_metrics() 

ASEs %>% 
  group_by(Model) %>% 
  summarise(ASE_mean = mean(ASE),
            ASE_median = median(ASE),
            ASE_sd = sd(ASE),
            num_batches = n())

```


```{r}
mdl_compare$get_tabular_metrics(ases = FALSE) %>% 
  na.omit()
```

### Remove unwanted models
```{r}
mdls_to_keep = c("VARS AIC Both B", "VARS AIC None B")
mdl_compare$keep_models(mdl_names = mdls_to_keep)
# mdls_to_remove = c("VARS AIC Const B", "VARS AIC Trend B", "VARS BIC Both B", "VARS BIC Const B", "VARS BIC None B", "VARS BIC Trend B")
# mdl_compare$remove_models(mdl_names = mdls_to_remove)
```

```{r}
models = list(
  "VARS AIC Both A" = list(select = "aic", trend_type = "both", lag.max = lag.max, sliding_ase = FALSE),
  "VARS AIC None A" = list(select = "aic", trend_type = "none", lag.max = lag.max, sliding_ase = FALSE),
  "VARS AIC Both B" = list(select = "aic", trend_type = "both", lag.max = lag.max, sliding_ase = TRUE),
  "VARS AIC None B" = list(select = "aic", trend_type = "none", lag.max = lag.max, sliding_ase = TRUE)
)

```


```{r}
mdl_compare = ModelCompareMultivariateVAR$new(data = all_data, mdl_list = models, var_interest = var_interest,
                                              n.ahead = n.ahead, batch_size = batch_size, verbose = 0)
```

```{r}
ASEs = mdl_compare$get_tabular_metrics() 

ASEs %>% 
  group_by(Model) %>% 
  summarise(ASE_mean = mean(ASE),
            ASE_median = median(ASE),
            ASE_sd = sd(ASE),
            num_batches = n())

```

### Simple Forecasts
```{r fig.width=10}
mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
mdl_compare$plot_simple_forecasts(lastn = FALSE, limits = TRUE)
```


### Batch Forecasts
```{r fig.width=10}
mdl_compare$plot_batch_forecasts(only_sliding = FALSE)
```

### Batch ASEs
```{r fig.width=10}
mdl_compare$plot_batch_ases(only_sliding = FALSE)
```

### ASE Histograms
```{r}
mdl_compare$plot_histogram_ases()  # Same function as univariate
```


### Statistical Compare
```{r}
mdl_compare$statistical_compare() 
```


# For Pre-live Manual Method
```{r}
plot(data)
```

* There seems to be some weekly trend in the data (downward)

## Train Test Split


```{r}
n = nrow(data)
train_data = all_data %>% dplyr::filter(row_number() <= (n - n.ahead))
test_data = all_data %>% dplyr::filter(row_number() > (n - n.ahead))
```


## VARselect
```{r}
# VARselect will select the best K for this VAR model
# Looks at the cross correlation structure (ccf) to figure this out 
vselect = VARselect(all_data, lag.max = lag.max, type = "both", season = NULL, exogen = NULL)
vselect # Gives AIC values for various K values

k = vselect$selection[["AIC(n)"]]  # BIC = SC(n)
k
```

AIC picks a VAR(10) model

## VAR Model

```{r}
lsfit = VAR(train_data, p=k, type="both")
stats::AIC(lsfit)
a = summary(lsfit)
#a
a$varresult[var_interest]  # $coefficients[,"Pr(>|t|)"]
```

## Predictions

```{r}
preds = stats::predict(lsfit, n.ahead=n.ahead)

results = preds$fcst[[var_interest]] %>% 
  dplyr::as_tibble() %>% 
  dplyr::mutate(Week = seq(n-n.ahead+1,n,1)) 

results

```

### ASE

```{r}
ASE_data = data %>% 
  dplyr::full_join(results, by = "Week") %>% 
  na.omit()

ASE_data
```

```{r}
ASE = mean((ASE_data$cmort - ASE_data$fcst)^2, na.rm = TRUE)
ASE
```


### Plots

```{r}
results = results %>% 
  dplyr::mutate(cmort = fcst) %>% 
  dplyr::mutate(Model = 'VAR Model') %>% 
  dplyr::select(Model, Week, cmort)

combined_data = data %>% 
  mutate(Model = 'Realization') %>% 
  dplyr::select(Model, Week, cmort ) %>% 
  bind_rows(results) %>% 
  mutate(Model = as.factor(Model))

combined_data
```


```{r fig.width=10}
p = ggplot2::ggplot() + 
  ggplot2::geom_line(data = combined_data, mapping = ggplot2::aes(x = Week, y = cmort, color = Model))

print(p)

plot(seq(1,n,1), data$cmort, type = "b")
points(seq(n-n.ahead+1,n,1), preds$fcst$cmort[,'fcst'], type = "b", pch = 15)
```


```{r fig.width=10}
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues")) # Change color pallet to make distinguishable. 
```


# Sales Read Data
```{r}
data = read.csv("../Datasets/businesssales.csv")
data %>% glimpse()
data %>%  skimr::skim()
```

```{r fig.width=10}
df.m = reshape::melt(data, id.vars = "X")
ggplot(df.m) + 
  facet_wrap(variable ~ ., ncol = 1, scales = 'free_y') + 
  geom_line(aes(x = X, y = value, colour = variable))

```

```{r}
all_data = data %>% dplyr::select(-X)
```

## CCF Analysis

```{r}
lag.max = 10
```

```{r}

var_interest = 'sales'

for (name in colnames(all_data)){
  # By convention, the X2 variable comes first
  c = ccf(data[name], data[var_interest], lag.max = lag.max)
  index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
  max_ccf = c$lag[,1,1][index] 
  cat(paste("\nFor variable: ", name, " , max cross-correlation is at lag: ", max_ccf, sep = ""))
}


```


## Multivariate Compare Object
```{r}

lag.max = 7

models = list(
  "VARS AIC Both A" = list(select = "aic", trend_type = "both", lag.max = lag.max, sliding_ase = FALSE),
  "VARS AIC Trend A" = list(select = "aic", trend_type = "trend", lag.max = lag.max, sliding_ase = FALSE),
  "VARS AIC Const A" = list(select = "aic", trend_type = "const", lag.max = lag.max, sliding_ase = FALSE),
  "VARS AIC None A" = list(select = "aic", trend_type = "none", lag.max = lag.max, sliding_ase = FALSE),
  "VARS BIC Both A" = list(select = "bic", trend_type = "both", lag.max = lag.max, sliding_ase = FALSE),
  "VARS BIC Trend A" = list(select = "bic", trend_type = "trend", lag.max = lag.max, sliding_ase = FALSE),
  "VARS BIC Const A" = list(select = "bic", trend_type = "const", lag.max = lag.max, sliding_ase = FALSE),
  "VARS BIC None A" = list(select = "bic", trend_type = "none", lag.max = lag.max, sliding_ase = FALSE)
)
```

```{r}
n.ahead = 5
```


```{r}
mdl_compare = ModelCompareMultivariateVAR$new(data = all_data, mdl_list = models, var_interest = var_interest,
                                              n.ahead = n.ahead, batch_size = batch_size, verbose = 1)
```

### AIC BIC Full Data
```{r}
mdl_compare$get_xIC()
mdl_compare$get_xIC(sort_by = "BIC")
```

### Tabular Metrics

```{r}
ASEs = mdl_compare$get_tabular_metrics() 

ASEs %>% 
  group_by(Model) %>% 
  summarise(ASE_mean = mean(ASE),
            ASE_median = median(ASE),
            ASE_sd = sd(ASE),
            num_batches = n()) %>% 
  dplyr::arrange((ASE_mean))

```

* Using AIC, "VARS AIC Const A" gets selected
* Using BIC, "VARS BIC Const A" gets selected
* Using ASE values also, these 2 models do the best
* We will keep these 2 models

### Keep limited models
```{r}
mdls_to_keep = c("VARS AIC Const A", "VARS BIC Const A")
mdl_compare$keep_models(mdl_names = mdls_to_keep)
```

### Comparing Forecasts
```{r}
forecasts = mdl_compare$get_tabular_metrics(ases = FALSE) 

forecasts %>% 
  na.omit()
```

```{r}
n = nrow(data)
test_data = data %>% dplyr::filter(row_number() > (n - n.ahead))
test_data
```

## MLR with correlated errors

```{r}
data2 = data %>% 
  mutate(ad_tv1 = dplyr::lag(ad_tv, 1)) %>% 
  mutate(ad_online1 = dplyr::lag(ad_online, 1)) %>% 
  mutate(discount3 = dplyr::lag(discount, 3)) 

data2
```

### Train Test Split
```{r}
n = nrow(data2)
train_data = data2 %>% dplyr::filter(row_number() <= (n - n.ahead))
test_data = data2 %>% dplyr::filter(row_number() > (n - n.ahead))
```

### Fit the model
```{r}
# Includes trend + original variables + lagged variables
ksfit = lm(sales ~ ., data = train_data)
aic_fit = aic.wge(ksfit$residuals,p=0:8, q=0)  
aic_fit
```

```{r}
fit = arima(train_data$sales, order=c(aic_fit$p,0,0), xreg=train_data %>% dplyr::select(-sales))
fit
```

### Evaluate Residuals
```{r}
tswgewrapped::evaluate_residuals(fit$residuals %>% na.omit())
```

### Compare Forecasts
```{r}
test_data2 = test_data %>% 
  dplyr::select(-sales)

test_data2
```

```{r}
preds = stats::predict(fit, newxreg = test_data2)
preds
test_data$sales
```

```{r}
ASE = mean((preds$pred - test_data$sales)^2)
ASE
```


```{r}


```
