---
title: "Time Series - Async Week 10"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
library(tswge)
library(tidyverse)
library(tseries)
library(orcutt)
```




# ARIMA - Box Jenkins ID

## 10.3 

### 3

```{r}
x = read.csv("../Datasets/10_year_bond_rate_2010-2015.csv")
x %>%  glimpse()
```

```{r}
x = x$Adj.Close
```

```{r}
check_stationarity(x)
```

### 4
```{r}
xdif1 = artrans.wge(x, phi.tr = 1)
px = plotts.sample.wge(xdif1)
```

### 5
```{r}
aic5.wge(xdif1)
```

### 6
```{r}
aic5.wge(x)
```

### 7

```{r}
aic.wge(x)
```

## 10.7

### 3

```{r}
e = est.arma.wge(x, p = 6)
```

### 5
```{r}
x = read.csv("../Datasets/zero_one_or_tworootsofone.csv")
x %>%  glimpse()
```

```{r}
x = x$x
```

```{r}
px = plotts.sample.wge(x)
```

* At least 1 unit root

```{r}
xdif1 = artrans.wge(x, phi.tr = 1)
px = plotts.sample.wge(xdif1)
```

* Seems like there is more wandering and extended ACF. May need to difference again
```{r}
xdif2 = artrans.wge(xdif1, phi.tr = 1)
px = plotts.sample.wge(xdif2)
```

Same thing using overfit table
```{r}
e = est.ar.wge(x, p = 8, type = 'burg')
```

Shows 2 roots close to unit circle.

## ADF (Augmented Dickey Fuller Test)
```{r}
phis = c(0.9, 0.95, 0.975)
ns = c(100, 200)
repeats = 10


for (phi in phis){
  for(n in ns){
    print(paste("Phi = ", phi, " n = ", n))
    for (i in 1:repeats){
      x = gen.arma.wge(n = n, phi = phi, plot = FALSE)    
      adf = adf.test(x)
      
      if (adf$p.value < 0.05){
        print("Reject Null!! Reject the null hypothesis that there is a non stationary model")
        #print("Reject the null hypothesis that there is a stationary model, i.e. model is non stationary")
      }
      else{
        print("Fail To Reject Null!! FTR the null hypothesis that there is a non stationary term")
        
      }
    }
  }
}

```

## 10.8

### 3
```{r}
x = gen.arma.wge(200,phi = c(.9), sn = 5)
adf.test(x)
```

### 4
```{r}
phis = c(0.9, 0.95, 0.99)
ns = c(200)
repeats = 10


for (phi in phis){
  for(n in ns){
    cat(paste("\n\nPhi = ", phi, " n = ", n))
    for (i in 1:repeats){
      x = gen.arma.wge(n = n, phi = phi, plot = FALSE)    
      adf = adf.test(x)
      
      if (adf$p.value < 0.05){
        cat("\nReject Null!! Reject the null hypothesis that there is a non stationary model")
        #print("Reject the null hypothesis that there is a stationary model, i.e. model is non stationary")
      }
      else{
        cat("\nFail To Reject Null!! FTR the null hypothesis that there is a non stationary term")
        
      }
    }
  }
}

```

## 10.9

### 4

```{r}
data = read.csv("../Datasets/swadelay.csv")
data %>% glimpse()
```

```{r}
x = data$arr_delay
```

```{r}
e.burg = est.ar.wge(x, p = 15, type = 'burg')
```

```{r}
factor.wge.season(12)
```

## 10.10

### 3
```{r}
data(airlog)
airlog %>% glimpse()
x = airlog
```

```{r}
px = plotts.sample.wge(x)
```

```{r}
d1 = artrans.wge(x, phi.tr = 1)
d1.12 = artrans.wge(d1, phi.tr = c(rep(0,11),1))
```

```{r}
e.burg = est.ar.wge(d1.12, p = 12, type = 'burg')
```

```{r}
psi.weights.wge(phi = e.burg$phi, lag.max = 10)
```

## 10.12

### 2
```{r}
x = gen.sigplusnoise.wge(100, b0 = 0, b1= 0, phi= .95, sn = 28)
```

```{r}
t = seq(1,100,1)
df = data.frame(x = x, t= t)
fit = lm(x~t, data = df)
summary(fit)
```

### 5

```{r}
phis = c(0.8, 0.95, 0.99)
ns = c(100, 200)
repeats = 10

results = tribble(~phi, ~n, ~total, ~num_rejects) # With Rolling Window

for (phi in phis){
  for(n in ns){
    cat(paste("\n\nPhi = ", phi, " n = ", n))
    count = 0
    for (i in 1:repeats){
      x = gen.sigplusnoise.wge(n = n, b0 = 0, b1 = 0, phi = phi, plot = FALSE)    

      t = seq(1,100,1)
      df = data.frame(x = x, t= t)
      fit = lm(x~t, data = df)
      #summary(fit)
      
      p_val_time = summary(fit)$coefficients[,4][2]

      if (p_val_time < 0.05){
        cat("\nReject Null!! Reject the null hypothesis that there is no time trend")
        count = count + 1
      }
      else{
        cat("\nFail To Reject Null!! FTR the null hypothesis that there is no time trend")
      }
    }
    
    results = results %>% add_row(phi = phi, n = n, total = repeats, num_rejects = count) 
  }
}

results

```

## Cochrane-Orcutt Test
```{r}
x = gen.sigplusnoise.wge(100, b0 = 0, b1= 0, phi= .95, sn = 21)  # No Signal b0 = 0, b1 = 0
```

```{r}
t = seq(1,100,1)
df = data.frame(x = x, t= t)
fit = lm(x~t, data = df)
summary(fit)
```

* This indicates a very significant slope, but that is a type 1 error.
* We will run Cochrane-Orcutt method to test for signal

```{r}
cfit = cochrane.orcutt(fit) 
summary(cfit)$coefficients[2,4]
```

* This p-value fails to reject the null hypothesis (right decision)

### Mini-Simulation

```{r}
phis = c(0.8, 0.95, 0.99)
ns = c(100, 200)
repeats = 10

results = tribble(~phi, ~n, ~total, ~num_rejects_org, ~num_rejects_co) # With Rolling Window

for (phi in phis){
  for(n in ns){
    cat(paste("\n\nPhi = ", phi, " n = ", n))
    count_org = 0
    count_co = 0
    for (i in 1:repeats){
      x = gen.sigplusnoise.wge(n = n, b0 = 0, b1 = 0, phi = phi, plot = FALSE)    

      t = seq(1,100,1)
      df = data.frame(x = x, t= t)
      fit = lm(x~t, data = df)
      #summary(fit)
      
      p_val_time_org = summary(fit)$coefficients[2,4]
      cfit = cochrane.orcutt(fit, max.iter = 500) 
      p_val_time_co = summary(cfit)$coefficients[2,4]
      
      if (p_val_time_org < 0.05){
        cat("\nReject Null!! Reject the null hypothesis that there is no time trend")
        count_org = count_org + 1
      }
      else{
        cat("\nFail To Reject Null!! FTR the null hypothesis that there is no time trend")
      }
      
      if (p_val_time_co < 0.05){
        cat("\nReject Null!! Reject the null hypothesis that there is no time trend")
        count_co = count_co + 1
      }
      else{
        cat("\nFail To Reject Null!! FTR the null hypothesis that there is no time trend")
      }
    }
    
    results = results %>% add_row(phi = phi, n = n, total = repeats, num_rejects_org = count_org, num_rejects_co = count_co) 
  }
}

results
```

## 10.13

### 3

```{r}
data = read.csv("../Datasets/swadelay.csv")
data %>% glimpse()
```

```{r}
x = data$arr_delay
```


```{r}
t = seq(1,length(x),1)
df = data.frame(x = x, t= t)
fit = lm(x~t, data = df)
summary(fit)
```

## 5 
```{r}
cfit = cochrane.orcutt(fit, max.iter = 500) 
summary(cfit)
```

```{r}

```

```{r}

```

```{r}

```

