---
title: "Time Series - Async Week 6"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tswge)
```

# Signal + Noise Models
```{r}
?gen.sigplusnoise.wge
```

## Linear Signal + White Noise
```{r}
xt = gen.sigplusnoise.wge(n = 100,
                          b0 = 2, b1 = 4, # Linear Trend
                          vara = 100, # White Noise
                          sn = 101)  
```

## Just White Noise (No Signal)
```{r}
xt = gen.sigplusnoise.wge(n = 100,
                          b0 = 0, b1 = 0, # No Trend (signal)
                          vara = 10, # White Noise
                          sn = 101) 
```

## Just AR Noise (No Signal)
```{r}
xt = gen.sigplusnoise.wge(n = 100,
                          b0 = 0, b1 = 0, # No Trend (signal)
                          phi = 0.975, vara = 10, # AR(1) Noise
                          sn = 101)
# Note that the wandering behavior can create a realization that looks like trend even though it is not trend.
```

## Periodic Signal
```{r}
# Xt = 5*cos(2*pi*0.1*t + 0.25) + Zt where Zt ~ AR(1)
xt = gen.sigplusnoise.wge(n = 100,
                          coef = c(5,0), freq = c(0.1, 0), psi = c(0.25,0), # Periodic Signal
                          phi = 0.975, vara = 4, # AR(1) Noise
                          sn = 101)
```

# AR(4) model compared to Global Temperature
```{r}
# Model in factored form: (1 - 0.975B)(1 - 0.2B + 0.45B2)(1 + 0.53B)*Xt = at
# Note that you feed in the phi values (note the sign)
parms = mult.wge(c(0.975), c(0.2,-0.45), c(-0.53))
parms$model.coef

# Check
factor.wge(phi = parms$model.coef)

#Now that we have the coefficients, () we can generate a realization.
xt = gen.arma.wge(160,
                  phi = parms$model.coef, vara = 1, # AR(4) Noise term
                  sn = 101)
```

# ARIMA Models

## Realization
```{r}
# ARIMA(0,1,0)
x1=gen.arima.wge(n = 200, d = 1)
px = plotts.sample.wge(x1)

# ARIMA(2,1,0)
x2=gen.arima.wge(n = 200, phi = c(1.5, -0.8), d = 1)
px = plotts.sample.wge(x2)

# ARIMA(2,2,1)
x3=gen.arima.wge(n = 200, phi = c(1.5,-0.8), d = 2, theta = -0.8)
px = plotts.sample.wge(x3)

```

## Make Stationary (Differencing)

```{r}

# the following command differences the data in x
# y = artrans.wge(x, phi.tr = 1)
# This simply means that y(i) = x(i) – x(i-1)
# y has length n-1 because x(1) has no x(0) before it.
# Example

x = c(1,3,6,10,25)
y = artrans.wge(x, phi.tr = 1)
y # shows the 4 differences
```

# Seasonal Models (ARUMA)

## Realization (Quarterly)

```{r}
# (1 - B^4)Xt = at
x1 = gen.aruma.wge(n = 80, s = 4, sn = 6)
px = plotts.sample.wge(x1)
# ACF repeats every 4 (4, 8, 12, 16)
# Spectral Densitiy shows peaks at f = 0, 0.25 and 0.5

# (1-B+0.6B^2)(1 - B^4)Xt = (1+0.5B)at
phi = c(1,-.6)
theta = -0.5
factor.wge(phi = phi)
factor.wge(phi = theta)

x2 = gen.aruma.wge(n = 80, phi = phi, s = 4, theta = theta, sn = 6)
px = plotts.sample.wge(x2)
# ACF repeats every 4 (4, 8, 12, 16)
# Spectral Densitiy shows peaks at f = 0, 0.25 but not so much at 0.5
# since this is getting cancelled by the MA term
# Also previously, there was a low value at 0.13, but now due to the 
# AR component, there is a much hgher value at this frequency.
```

## Realization (Monthly)

```{r}

# (1 - B^12)Xt = at
x3 = gen.aruma.wge(n = 180, s = 12, sn = 6)
px = plotts.sample.wge(x3,lag.max = 48)
# ACF repeats every 12 (12, 24, 36, 48)
# Spectral Densitiy shows 5 peaks between f = 0 and 0.5

# (1-B+0.6B^2)(1 - B^12)Xt = (1+0.5B)at
x4 = gen.aruma.wge(n = 180, phi = phi, s = 12, theta = theta, sn = 6)
px = plotts.sample.wge(x4,lag.max = 48)
# ACF repeats every 12 (12, 24, 36, 48)
# Spectral Densitiy shows 5 peaks between f = 0 and 0.5.
# However, the dB value at f = 0.13 has increased due to the peak from the AR term
# and at f = 0.5 has decreased due to te dip from the MA term

```

## Make Stationary (Differencing)

### Quarterly
```{r}
x = gen.aruma.wge(n = 80, s = 4, sn = 81) #tswge function to generate ARIMA and Seasonal Models
dif = artrans.wge(x = x, phi.tr = c(0,0,0,1)) #Take out the (1-B^4)
aic5.wge(dif) #Check the structure of the noise
```

### Annual (not monthly)
```{r}
phi = c(.4,.6,-.74)
theta = c(-.7)

factor.wge(phi = phi)
factor.wge(phi = theta)

x = gen.aruma.wge(n = 80, phi = phi, theta = theta, s=12, sn = 31)
dif = artrans.wge(x = x, phi.tr = c(rep(0,11),1)) #Take out the (1-B^12)
aic5.wge(dif) #Check the structure of the noise
```

## Factoring
```{r}
# Factor Table
s = 4; factor.wge(c(rep(0,s-1), 1))   # Quarterly
s = 12; factor.wge(c(rep(0,s-1), 1))  # Monthly
s = 5; factor.wge(c(rep(0,s-1), 1))   # Seasonality every 5 time periods


# factoring 1
phi = c(-0.2, 0.4, 0.49, 0, 0, 0, 0, 0, 0, 0, 0, 1, 0.2, -0.4, -0.49)
factor.wge(phi = phi)

# factoring 2
phi = c(-0.3, 0.8, 0 , 0, 1, 0.3, -0.8)
factor.wge(phi = phi)

```

# More General Seasonal Model

## Airline Data
```{r}
data(airlog)
px = plotts.sample.wge(airlog)
# Slowly decaying ACF --> ARIMA
# Spectral Densirty shows yearly seasonality (5 peaks between f = 0 and f = 0.5)
```


```{r}
# take first differences of the data
SA1 = artrans.wge(x = airlog, phi.tr = 1) 

# take the 12th difference of the data (1-B12)
SA12 = artrans.wge(x = airlog, phi.tr = c(rep(0,11),1))

# take the 12th difference of the first difference (1-B)(1-B12)
SA1_12 = artrans.wge(x = SA1, phi.tr = c(rep(0,11),1)) 
```

```{r}
Parzen = aic.wge(SA12, p = 13) # Phi(B)(1-B12)(Xt-mu) = at
Box = aic.wge(SA1_12, q = 13) # (1-B)(1-B12)(Xt-mu) = Theta(B)at
WoodwardAndGray = aic.wge(SA1_12, p = 12) # Phi(B)(1-B)(1-B12)(Xt-mu) = at

Parzen$value
Box$value
WoodwardAndGray$value

# We will return to this competition when we use the model to forecast!
```

```{r}
Nikhil = aic5.wge(SA1_12, p = 0:15, q = 0:15)
Nikhil  # Matches top one from Dr. Box
```

# Async

```{r}
phi = c(0.6, -0.8)
theta = -0.3
xt = gen.arima.wge(n = 500, phi = phi, d = 2, theta = theta, sn = 37)

# 2
factor.wge(phi = phi)

# 3
px = parzen.wge(xt)
```

```{r}
phi = c(0.6, -0.8)
theta = -0.3
xt = gen.arima.wge(n = 500, phi = phi, d = 2, theta = theta, sn = 35)

first_diff = artrans.wge(xt, phi.tr = 1)
second_diff = artrans.wge(first_diff, phi.tr = 1)

aic5.wge(second_diff)

```

```{r}
phi = c(0.6, -0.94)
theta = -0.3
xs = gen.aruma.wge(n = 200, phi = phi, s = 6, theta = theta, sn = 19)
px = plotts.sample.wge(xs)
```

```{r}
phi = c(0.6, -0.8)
theta = c(0.3, -0.7)
s = 12

xt = gen.aruma.wge(n = 500, phi = phi, theta = theta, s = s, sn = 37)
dif = artrans.wge(x = xt, phi.tr = c(rep(0,s-1),1))
# dif1 = artrans.wge(x = dif, phi = 1)
# dif2 = artrans.wge(x = dif1, phi = 1)
aic5.wge(x = dif)
```

```{r}
phi = c(-0.5, 0.2, 0, 1, 0.5, -0.2)
factor.wge(phi = phi)

phi = c(-0.3, 1.2, 0.4, 0, 0.5, 0, 0, 0, 0 ,0, 0, 1, 0.3, -1.2, -0.4)
factor.wge(phi = phi)
```


```{r}

```

