---
title: "Time Series - Async Week 12"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
library(tswge)
library(tswgewrapped)
library(tidyverse)
library(tseries)
library(orcutt)
library(vars)
library(RColorBrewer)
```

# Multiple Linear Regression with Correlated Errors 

## Sales Data

### Basic Model (no time trend)
```{r}
#Assuming a data.frame exists with corresponding names below.
BSales = read.csv("../Datasets/businesssales.csv", header = TRUE)
BSales %>%  glimpse()
```

```{r}
# All data with no lag and no trend
ksfit = lm(sales ~ ad_tv + ad_online + discount, data = BSales)
aic_fit = aic.wge(ksfit$residuals,p=0:8, q=0)  # AIC picks p=7

# These values will be very different from the ones obtained from arima fit below
ksfit
aic_fit
```

```{r}
fit=arima(BSales$sales,order=c(aic_fit$p,0,0),xreg=BSales[,3:5])
fit
```

* ad_tv, online_ad and discount does not seem to be significant

```{r}
tswgewrapped::evaluate_residuals(fit$residuals, model_name = "Multivar with corr errors", lag.max = 20)
```

* **Residuals are not white noise**

### Basic Model (with time trend)

```{r}
# All data with no lag and no trend
ksfit = lm(sales ~ X + ad_tv + ad_online + discount, data = BSales)
aic_fit = aic.wge(ksfit$residuals,p=0:8, q=0)  # AIC picks p=7

# These values will be very different from the ones obtained from arima fit below
ksfit
aic_fit
```

```{r}
fit = arima(BSales$sales, order=c(aic_fit$p,0,0), xreg=BSales[,c(1,3:5)])
fit
```

* Time (X) is significant, other variables are not

```{r}
tswgewrapped::evaluate_residuals(fit$residuals)
```

* Still not white noise

### Using Lagged Variables

#### Creating lagged variables
```{r}
df = data.frame(Y = c(1,1,2,3,4,4,5,8),X1 = c(5,6,6,7,7,8,8,9))

df = df %>% 
  mutate(X1_L1 = dplyr::lag(X1, 1)) %>% 
  mutate(X1_L2 = dplyr::lag(X1, 2)) 

df
```

#### 12.2 (5)
```{r}
BSales = BSales %>% 
  mutate(ad_tv1 = dplyr::lag(ad_tv, 1)) %>% 
  mutate(ad_online1 = dplyr::lag(ad_online, 1)) 

BSales

```

#### Sales Model with lagged variables

```{r}
pairs(BSales)

data = BSales %>% 
  dplyr::select(sales, ad_tv1, ad_online1, discount)
```


```{r}
ksfit = lm(sales ~ ad_tv1 + ad_online1 + discount, data = data)
aic_fit = aic.wge(ksfit$residuals,p=0:8, q=0)  # AIC picks p=7
```

```{r}
fit = arima(data$sales, order=c(aic_fit$p,0,0), xreg=data %>% dplyr::select(-sales))
fit
```

* Now ad_tv1 and ad_online1 are quite significant
* discount is also coming out to be slightly significant now

#### Sales Model with trend + lagged variables
```{r}
data = BSales %>% 
  dplyr::select(sales, X, ad_tv1, ad_online1, discount)  
```

```{r}
ksfit = lm(sales ~ ., data = data)
aic_fit = aic.wge(ksfit$residuals,p=0:8, q=0)  # AIC picks p=7
```

```{r}
fit = arima(data$sales, order=c(aic_fit$p,0,0), xreg=data %>% dplyr::select(-sales))
fit
```

```{r}
tswgewrapped::evaluate_residuals(fit$residuals %>% na.omit())
```

* Residuals look like white noise.

#### Finding what lag to use

```{r}
# By convention, the X2 variable comes first
c = ccf(BSales$ad_tv, BSales$sales, lag.max = 10)
index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
max_ccf = c$lag[,1,1][index] 
max_ccf
```

```{r}
c = ccf(BSales$ad_online, BSales$sales, lag.max = 10)
index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
max_ccf = c$lag[,1,1][index] 
max_ccf
```


```{r}
c = ccf(BSales$discount, BSales$sales, lag.max = 10)
index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
max_ccf = c$lag[,1,1][index] 
max_ccf
```

#### 12.3 (4)
```{r}
data = read.csv("../Datasets/whatisthelag.csv")
data %>% glimpse()
```

```{r}
c = ccf(data$X1, data$Y, lag.max = 10)
index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
max_ccf = c$lag[,1,1][index] 
max_ccf
```

# VAR Models

## Basic Example

```{r}
x1.25=c( -1.03,  0.11, -0.18, 0.20, -0.99, -1.63, 1.07,  2.26, -0.49, -1.54,  0.45,  0.92,
         -0.05, -1.18,  0.90,  1.17,  0.31,  1.19,  0.27, -0.09,  0.23, -1.91,  0.46,  3.61, -0.03)
x2.25=c( -0.82,  0.54,  1.13, -0.24, -0.77,  0.22,  0.46, -0.03, -0.59,  0.45,  0.59,  0.15,
         0.60,  0.13, -0.04,  0.12, -0.96,  0.23,  1.81, -0.01, -0.95, -0.55, -0.15,  0.71,  0.90)


px = plotts.wge(x1.25)
px = plotts.wge(x2.25)
```


```{r}
# Only use first 20 to forecast the next 5
x1=x1.25[1:20]
x2=x2.25[1:20] 
```

### Univariate Analysis
```{r}
p1=aic.wge(x1,p=0:8,q=0:0)
# aic picks p=2
x1.est=est.ar.wge(x1,p=p1$p)
fore.arma.wge(x1,phi=x1.est$phi,n.ahead=5,lastn=FALSE,limits=FALSE)
```

```{r}
p2=aic.wge(x2,p=0:8,q=0:0)
# aic picks p=2
x2.est=est.ar.wge(x1,p=p2$p)
fore.arma.wge(x2,phi=x2.est$phi,n.ahead=5,lastn=FALSE,limits=FALSE)
```

### Bivariate Analysis

* Can X1 tell us something about X2 and vice versa

```{r}
# By convention, the X2 variable comes first
c = ccf(x2.25, x1.25, lag.max = 10)
index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
max_ccf = c$lag[,1,1][index] 
max_ccf
```

* Lag -5 for x2 seems to be important to make a prediction for x1

```{r}
X = cbind(x1,x2)
X %>% head()
```


```{r}
# VARselect will select the best K for this VAR model
# Looks at the cross correlation structure (ccf) to figure this out 
vselect = VARselect(X, lag.max = 6, type = "const", season = NULL, exogen = NULL)
vselect # Gives AIC values for various K values

k = vselect$selection[["AIC(n)"]]
k
#VARselect picks p=5 (using AIC)
```


```{r}
lsfit = VAR(X, p=k, type="const")
summary(lsfit)  
```

```{r}
preds = predict(lsfit, n.ahead=5)
preds$fcst
# preds$fcst$x1[,1]
# preds$fcst$x1[1,1] - [5,1] are the VAR forecasts for x1.  Similar for x2
```


```{r}
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues")) # Change color pallet to make distinguishable. 
```

* Upper and Lower Confidence intervals for X1 are almost the same

# Melanoma Data

```{r}
# melanoma incidence and sunspot numbers 1936-1972
melanoma=c(1.0, 0.9, 0.8, 1.4, 1.2, 1.0, 1.5, 1.9, 1.5, 1.5, 1.5, 1.6, 1.8, 2.8, 2.5, 2.5, 2.4, 2.1, 1.9, 2.4, 2.4, 2.6, 2.6, 4.4, 4.2, 3.8, 3.4, 3.6, 4.1, 3.7, 4.2, 4.1, 4.1, 4.0, 5.2, 5.3, 5.3)
sunspot=c(40, 115, 100,  80,  60,  40,  23,  10,  10,  25,  75, 145, 130, 130,  80,  65,  20,  10,   5,  10, 60, 190, 180, 175, 120,  50,  35,  20,  10,  15,  30,  60, 105, 105, 105,  80,  65)
```

## VAR Model

```{r}
# Use data till 1967 to make prediction from 1968-1972 (5 years)
mel.67=melanoma[1:32]
sun.67=sunspot[1:32] 

mel.67=melanoma[1:29]
sun.67=sunspot[1:29] 
```

### Univariate Analysis
```{r}
p.mel=aic.wge(mel.67,p=0:8,q=0:0)
p.mel$p
mel.est=est.ar.wge(mel.67,p=p.mel$p)
f = fore.arma.wge(mel.67,phi=mel.est$phi,n.ahead=5,lastn=FALSE,limits=FALSE)
```

```{r}
p.sun=aic.wge(sun.67,p=0:8,q=0:0)
p.sun$p
sun.est=est.ar.wge(sun.67,p=p.sun$p)
f = fore.arma.wge(sun.67,phi=sun.est$phi,n.ahead=5,lastn=FALSE,limits=FALSE)
```


### Bivariate Analysis

```{r}
# By convention, the X2 variable comes first
c = ccf(sunspot, melanoma, lag.max = 12)
index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
max_ccf = c$lag[,1,1][index] 
max_ccf
```

```{r}
X = cbind(mel.67,sun.67)
X %>% head()
```

```{r}
# VARselect will select the best K for this VAR model
# Looks at the cross correlation structure (ccf) to figure this out 
vselect = VARselect(X, lag.max = 6, type = "const", season = NULL, exogen = NULL)
vselect # Gives AIC values for various K values

k = vselect$selection[["AIC(n)"]]
k
#VARselect picks p=4 (using AIC)
```


```{r}
lsfit = VAR(X, p=k, type="const")
summary(lsfit)  
```

```{r}
preds = predict(lsfit, n.ahead=8)
preds$fcst
```


```{r}
plot(seq(1,37,1),melanoma, type = "b", ylim = c(0,6))
points(seq(30,37,1), preds$fcst$mel.67[1:8,1], type = "b", pch = 15)

fanchart(preds, colors = brewer.pal(n = 8, name = "Blues")) # Change color pallet to make distinguishable. 
```

```{r}
ASE = mean((preds$fcst$mel.67[1:8,1] - melanoma[30:37])^2)
ASE
```

## Multiple Linear Regression with Correlated Errors

```{r}
sunspot
melanoma

sun.67
mel.67
```

```{r}
#data = data.frame(melanoma = mel.67, sunspot = sun.67, t = 1:32) %>% 
data = data.frame(melanoma = melanoma, sunspot = sunspot, t = 1:37) %>% 
  mutate(sunspot.2 = dplyr::lag(sunspot, 2)) %>% 
  dplyr::select(melanoma, sunspot.2, t)

train_data = data %>% 
  dplyr::filter(t <= 32)
```

```{r}
ksfit = lm(melanoma ~ ., data = train_data)
aic_fit = aic.wge(ksfit$residuals,p=0:8, q=0)  # AIC picks p=7

# These values will be very different from the ones obtained from arima fit below
ksfit
aic_fit
```

```{r}
fit=arima(train_data$melanoma, order=c(aic_fit$p,0,0),xreg= train_data %>% dplyr::select(-melanoma))
fit
```

```{r}
test_data = data %>% 
  dplyr::filter(t >= 33) %>% 
  dplyr::select(-melanoma)

test_data

preds = predict(fit, newxreg = test_data)  
plot(seq(1,37,1), melanoma, type = "b")
points(seq(33,37,1), preds$pred, type = "b", pch = 15)
```

```{r}
tswgewrapped::evaluate_residuals(fit$residuals %>% na.omit(), model_name = "Multivar with corr errors", lag.max = 50, k = 24)
```

## 12.5 

### 2

```{r}
#Assuming a data.frame exists with corresponding names below.
BSales = read.csv("../Datasets/businesssales.csv", header = TRUE)
BSales %>%  glimpse()
BSales %>% skimr::skim() # No Missing Values
```

```{r}
# VARselect will select the best K for this VAR model
# Looks at the cross correlation structure (ccf) to figure this out 
vselect = VARselect(BSales %>% dplyr::select(-X), lag.max = 6, type = "const", season = NULL, exogen = NULL)
vselect # Gives AIC values for various K values

k = vselect$selection[["SC(n)"]]  # BIC = SC(n)
k
```

### 3

#### With Trend (and constant)

```{r}
lsfit = VAR(BSales %>% dplyr::select(-X), p=k, type="trend")
stats::AIC(lsfit)
summary(lsfit)
```

#### Without Trend
```{r}
lsfit = VAR(BSales %>% dplyr::select(-X), p=k, type="const")
stats::AIC(lsfit)
summary(lsfit)  
```


# Backup

```{r eval=FALSE, include=FALSE}


# ASE for model with no lag and no trend (last 5)
BSales2 = BSales[1:95,]
ksfit=lm(sales~ad_tv+ad_online+discount, data = BSales2)
aic.wge(ksfit$residuals,p=0:8, q=0)  # AIC picks p=7
fit=arima(BSales2$sales,order=c(7,0,0),xreg=cbind(BSales2$ad_tv,BSales2$ad_online,BSales2$discount))
fit

preds = predict(fit, newxreg = cbind(BSales$ad_tv[96:100],BSales$ad_online[96:100],BSales$discount[96:100]))
ASE1 = mean((BSales$sales[96:100] - preds$pred)^2)
ASE1

dev.off()
plot(seq(1,100,1), BSales$sales[1:100], type = "l",xlim = c(0,100), ylab = "Business Sales", main = "5 Week Sales Forecast")
lines(seq(96,100,1), preds$pred, type = "l", col = "red")


# Model 2

#with trend

# ASE for model with no lag and trend (last 5)
t=1:100
BSales$t = t
BSales2 = BSales[1:95,]
ksfit=lm(sales~t+ad_tv+ad_online+discount, data = BSales2)
aic.wge(ksfit$residuals,p=0:8, q=0)  # AIC picks p=7
fit=arima(BSales2$sales,order=c(6,0,0),xreg=cbind(BSales2$ad_tv,BSales2$ad_online,BSales2$t,BSales2$discount))
fit

preds = predict(fit, newxreg = cbind(BSales$ad_tv[96:100],BSales$ad_online[96:100],BSales$t[96:100],BSales$discount[96:100]))
ASE2 = mean((BSales$sales[96:100] - preds$pred)^2)
ASE2

dev.off()
plot(seq(1,100,1), BSales$sales[1:100], type = "l",xlim = c(0,100), ylab = "Business Sales", main = "5 Week Sales Forecast")
lines(seq(96,100,1), preds$pred, type = "l", col = "red")


# Model 3

#Lagging BSales Ad Variables
ad_tv1 = dplyr::lag(BSales$ad_tv,1)
ad_online1 = dplyr::lag(BSales$ad_online,1)
BSales$ad_tv1= ad_tv1
BSales$ad_online1 = ad_online1


#with trend and lagging

# ASE for model with no lag and trend (last 5)
t=1:100
BSales$t = t
BSales2 = BSales[2:95,]
ksfit=lm(sales~t+ad_tv1+ad_online1+discount, data = BSales2)
aic.wge(ksfit$residuals,p=0:8,q=0:0)  # AIC picks p=7
fit = arima(BSales2$sales,order = c(7,0,0), xreg = cbind(BSales2$ad_tv1,BSales2$ad_online1,BSales2$t,BSales2$discount))
fit


preds = predict(fit, newxreg = cbind(BSales$ad_tv1[96:100],BSales$ad_online1[96:100],BSales$t[96:100],BSales$discount[96:100]))
ASE3 = mean((BSales$sales[96:100] - preds$pred)^2)
ASE3


dev.off()
plot(seq(1,100,1), BSales$sales[1:100], type = "l",xlim = c(0,100), ylab = "Business Sales", main = "5 Week Sales Forecast")
lines(seq(96,100,1), preds$pred, type = "l", col = "red")



####### Forecast Features 

plotts.sample.wge(BSales$ad_tv)
aic5.wge(BSales$ad_tv)
est_ad_tv = est.arma.wge(BSales$ad_tv,p = 2, q = 2)
ad_tvFORECAST = fore.arma.wge(BSales$ad_tv,phi = est_ad_tv$phi, theta = est_ad_tv$theta, n.ahead = 6)

plotts.sample.wge(BSales$ad_online)
aic5.wge(BSales$ad_online, p = 0:10)
est_online = est.arma.wge(BSales$ad_online,p = 6)
dev.off()
plot.ts(BSales$ad_online[1:100])
ad_onlineFORECAST = fore.arma.wge(BSales$ad_online,phi = est_online$phi, n.ahead = 6)



#with trend and lagging

ad_tvFORECAST1 = lag(ad_tvFORECAST,1)

ad_onlineFORECAST1 = lag(ad_onlineFORECAST,1)

# ASE for model with no lag and trend (last 5)
t=1:100
BSales$t = t
BSales2 = BSales[2:95,]
ksfit=lm(sales~t+ad_tv1+ad_online1, data = BSales2)
aic.wge(ksfit$residuals,p=0:8,q=0:0)  # AIC picks p=7
fit = arima(BSales2$sales,order = c(7,0,0), xreg = cbind(BSales2$ad_tv1,BSales2$ad_online1,BSales2$t))
fit


preds = predict(fit, newxreg = cbind(ad_tvFORECAST$f[2:6],ad_onlineFORECAST$f[2:6],BSales$t[96:100]))
ASE3.5 = mean((BSales$sales[96:100] - preds$pred[1:5])^2)
ASE3.5


plot(seq(1,100,1), BSales$sales[1:100], type = "l",xlim = c(0,100), ylab = "Business Sales", main = "5 Week Sales Forecast")
lines(seq(96,100,1), preds$pred[2:6], type = "l", col = "red")


```


```{r eval=FALSE, include=FALSE}
############ VAR MODELS ##########################

#Model 4

BSVar = VAR(cbind(BSales2$sales,BSales2$ad_tv1,BSales2$ad_online1), type = "both", lag.max = 10)

preds = predict(BSVar,n.ahead = 5)
                  
ASE4 = mean((BSales$sales[96:100] - preds$fcst$y1[,1])^2)
ASE4


dev.off()
plot(seq(1,100,1), BSales$sales[1:100], type = "l",xlim = c(0,100), ylab = "Business Sales", main = "5 Week Sales Forecast")
lines(seq(96,100,1), preds$fcst$y1[,1], type = "l", col = "red")


#Model 5
BSVar = VAR(cbind(BSales2$sales,BSales2$ad_tv,BSales2$ad_online), type = "both", lag.max = 10)

preds = predict(BSVar,n.ahead = 5)

ASE5 = mean((BSales$sales[96:100] - preds$fcst$y1[,1])^2)
ASE5

dev.off()
plot(seq(1,100,1), BSales$sales[1:100], type = "l",xlim = c(0,100), ylab = "Business Sales", main = "5 Week Sales Forecast")
lines(seq(96,100,1), preds$fcst$y1[,1], type = "l", col = "red")

```

