---
title: "Time Series - Async Week 11"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup
```{r}
library(tswge)
library(tswgewrapped)
library(tidyverse)
library(tseries)
library(orcutt)
```


# Ljung - Box Test for White Noise

## ARMA(2,1)
```{r}
x=gen.arma.wge(n=100, phi=c(1.6,-.9), theta=.8, sn=67)
x=x+10
px = plotts.sample.wge(x)
```

```{r}
aic.wge(x,p=0:8,q=0:4)
# AIC picks ARMA(2,1)
```

```{r}
x21=est.arma.wge(x,p=2,q=1)
# x21$phi: 1.6194830 -0.9131788
# x21$theta: 0.868127
# x21$vara: 1.076196
mean(x)  # 10.07557
```

### Testing White Noise Terms

#### Step 1: Visual Check
```{r}

# x21$res: Contains residuals from the ARMA(2,1) fit
px = plotts.wge(x21$res)
acf(x21$res)
```

#### Step 2: Ljung-Box Test

```{r}
ljung.wge(x21$res, p=2, q=1)
# $K: 24  (default)
# $chi.square: 20.92251
# $df: 21
# $pval: 0.4636851
```

```{r}
ljung.wge(x21$res,p=2,q=1,K=48)
# $K: 48 
# $chi.square: 44.93891
# $df: 45
# $pval: 0.4636851
```

## ARIMA with s = 12
```{r}
x=gen.aruma.wge(n=200,s=12,phi=c(1.5,-.8),sn=87)
x=x+50
px = plotts.sample.wge(x,lag.max=60)
```

```{r}
y=artrans.wge(x,phi.tr=c(0,0,0,0,0,0,0,0,0,0,0,1))
```

```{r}
aic.wge(y, type = 'bic')
```

```{r}
est.y = est.ar.wge(y,p=2)
```

### Testing White Noise Terms

#### Step 1: Visual Check
```{r}

# x21$res: Contains residuals from the ARMA(2,1) fit
px = plotts.wge(est.y$res)
acf(x21$res)
```

#### Step 2: Ljung-Box Test

```{r}
ljung.wge(est.y$res,p=2)
# $K: 24  (default)
# $chi.square: 20.95537
# $df: 22
# $pval: 0.5234977
```

```{r}
ljung.wge(est.y$res,p=2,K=48)
# $K: 48 
# $chi.square: 47.62509
# $df: 46
# $pval: 0.4063877
```

## Airlog
```{r}
data("airlog")
x = airlog
```

```{r}
px = plotts.sample.wge(x)
```

### Overfit Table


```{r}
factor.wge.season(12)
```


```{r}
e = est.ar.wge(x, p = 24, type = 'burg')
```

* Has (1-B) and (1-B12) terms

### Transformation
```{r}
# Difference the data
d1=artrans.wge(airlog,phi.tr=1)
```

```{r}
# Transform differenced data by 1-B^12
s12=c(0,0,0,0,0,0,0,0,0,0,0,1)
d1.12=artrans.wge(d1,phi.tr=s12)
```

### Fit Stationary
```{r}
aic.wge(d1.12,p=0:15,q=0:3)
# aic and aicc pick ARMA(12,1)
```

```{r}
# estimate parameters of stationary part
est.12.1=est.arma.wge(d1.12, p=12, q=1)
```

### Testing White Noise Terms

#### Step 1: Visual Check
```{r}
px = plotts.wge(est.12.1$res)
acf(est.12.1$res, lag.max = 50)
```

#### Step 2: Ljung-Box Test

```{r}
ljung.wge(est.12.1$res, p=12, q = 1)
```

```{r}
ljung.wge(est.12.1$res, p=12, q = 1, K=48)
```

## 11.3 #2

```{r}
x = airlog
```


### Transformation
```{r}
# Difference the data
d1=artrans.wge(airlog,phi.tr=1)
```

```{r}
# Transform differenced data by 1-B^12
s12=c(0,0,0,0,0,0,0,0,0,0,0,1)
d1.12=artrans.wge(d1,phi.tr=s12)
```

### Estimate Stationary Part
```{r}
# estimate parameters of stationary part
ww = est.arma.wge(d1.12, p=12)
```

### Testing White Noise Terms

#### Step 1: Visual Check
```{r}
px = plotts.wge(ww$res)
acf(ww$res)
```

#### Step 2: Ljung-Box Test

```{r}
ljung.wge(ww$res, p=12)
```

```{r}
ljung.wge(ww$res, p=12, K=48)

```

## Global Temperature Data

```{r}
data(hadley)
px = plotts.sample.wge(hadley)
```

### Startionary Model
```{r}
aic5.wge(hadley,p=0:6,q=0:1)
# AIC picks an ARMA(3,1) stationary model
```

```{r}
had.est=est.arma.wge(hadley,p=3,q=1)
# $phi: 1.2700171 -0.4685313  0.1911988
# $theta: 0.6322319
# $avar: 0.01074178

mean(hadley)  # -0.1684937
```

#### Checking White Noise Terms
```{r}
px = plotts.sample.wge(had.est$res, arlimits=TRUE)
```

```{r}
ljung.wge(had.est$res,p=3,q=1)
```

```{r}
ljung.wge(had.est$res,p=3,q=1,K=48)
```

#### Checking Similar Realizations
```{r}
px = gen.arma.wge(n = length(hadley), phi = had.est$phi, theta = had.est$theta, vara = had.est$avar)
px = gen.arma.wge(n = length(hadley), phi = had.est$phi, theta = had.est$theta, vara = had.est$avar)
px = gen.arma.wge(n = length(hadley), phi = had.est$phi, theta = had.est$theta, vara = had.est$avar)
px = gen.arma.wge(n = length(hadley), phi = had.est$phi, theta = had.est$theta, vara = had.est$avar)
```

* These realizations all have wandering behavior (although they show decreasing values in some)
* This may be consistent with the global temp data

### Non Stationary Model

#### Overfit Table
```{r}
e = est.ar.wge(hadley, p = 6)
```

* Consistent with (1-B) term

#### Dickey Fuller Test
```{r}
adf = adf.test(hadley)
adf$p.value
```

* ADF also points to a non stationary model

#### Transform Data
```{r}
d1.temp=artrans.wge(hadley,phi.tr=1)
px = plotts.sample.wge(d1.temp, arlimits=TRUE)
```

* Not really White noise since the 1st 2 are outside the limit lines.

```{r}
aic5.wge(d1.temp, p=0:6, q=0:1)
# AIC selects an ARMA(2,1)
```

```{r}
d1.temp.est = est.arma.wge(d1.temp, p=2, q=1)
# $phi: 0.3274341 -0.1786827
# $theta: 0.704618
# $avar: 0.01058826
```

#### Testing White Noise Terms
```{r}
px = plotts.sample.wge(d1.temp.est$res,arlimits=TRUE)
```

```{r}
ljung.wge(d1.temp.est$res,p=2,q=1)
```

```{r}
ljung.wge(d1.temp.est$res,p=2,q=1,K=48)
```

#### Checking Similar Realizations
```{r}
px = gen.aruma.wge(n = length(hadley), d = 1, phi = d1.temp.est$phi, theta = d1.temp.est$theta, vara = d1.temp.est$avar)
px = gen.aruma.wge(n = length(hadley), d = 1, phi = d1.temp.est$phi, theta = d1.temp.est$theta, vara = d1.temp.est$avar)
px = gen.aruma.wge(n = length(hadley), d = 1, phi = d1.temp.est$phi, theta = d1.temp.est$theta, vara = d1.temp.est$avar)
px = gen.aruma.wge(n = length(hadley), d = 1, phi = d1.temp.est$phi, theta = d1.temp.est$theta, vara = d1.temp.est$avar)
```

#### Forecast Comparison

```{r}
f = fore.arma.wge(hadley,phi=c(1.27,-.47,.19), theta=.63, n.ahead=50, limits=TRUE)
```

```{r}
f = fore.aruma.wge(hadley,d=1,phi=c(.33,-.18), theta=.7, n.ahead=50, limits=TRUE)
```

### Signal Plus Noise Model

#### Fit a GLM model
```{r}
x=hadley
n=length(x)
t=1:n
d = lm(x~t)
d
```

```{r}
# x.z are the residuals from the regression line
x.z = x-d$coefficients[1]-d$coefficients[2]*t 

# Could have also used d$residuals to compute residuals
r = d$residuals

all.equal(x.z, unname(r))
```

#### Manual Cochrane Orcutt Test
```{r}
# Fit AR(p) to the residuals
ar.z = aic.wge(x.z,p=0:6)
ar.z$p # is the order p
ar.z$phi # is vector of ar.z$p estimated AR coefficients
```

```{r}
# We transform the original data to get Yt = Phi(B)Xt
y.trans = artrans.wge(hadley, phi.tr=ar.z$phi)  
```

```{r}
# Transform the time variable
#ar.z$phi is vector of ar.z$p estimated AR coefficients
t.trans = artrans.wge(t, phi.tr=ar.z$phi)
```


```{r}
# For a linear Regression model to the transformed variables to check if the trend still exists after removing the autocorrelation terms.
fit = lm(y.trans~t.trans)
summary(fit)
```

#### Evaluate White Noise Terms
```{r}
plot(fit$residuals)
acf(fit$residuals)
```

```{r}
ljung.wge(fit$residuals, p = 4)  # In the videos p was not provided - why?
```

```{r}
ljung.wge(fit$residuals, p = 4, K = 48) # In the videos p was not provided - why?
```

#### Checking Similar Realizations
```{r}
x = gen.sigplusnoise.wge(length(hadley), b0 = -0.5257, b1 = 0.0044, phi = ar.z$phi, vara = 0.0103)
x = gen.sigplusnoise.wge(length(hadley), b0 = -0.5257, b1 = 0.0044, phi = ar.z$phi, vara = 0.0103)
x = gen.sigplusnoise.wge(length(hadley), b0 = -0.5257, b1 = 0.0044, phi = ar.z$phi, vara = 0.0103)
x = gen.sigplusnoise.wge(length(hadley), b0 = -0.5257, b1 = 0.0044, phi = ar.z$phi, vara = 0.0103)
```

* These loook consistent with the actual data

#### Forecast 
```{r}
f = fore.sigplusnoise.wge(hadley, linear = TRUE, max.p=4, n.ahead=50, limits=TRUE)
```

## Sunspot Data

```{r}
data("sunspot.classic")
p = plotts.sample.wge(sunspot.classic, lag.max = 50, arlimits = TRUE)
```

### Box-Jenkins Model
#### Model ID
```{r}
acf(sunspot.classic)
pacf(sunspot.classic)
```

* PACF is indicative of a AR(2) model

#### Parameter Estimation
```{r}
s2=est.ar.wge(sunspot.classic, p=2)
s2$phi # 1.3347282 -0.6474423
s2$avar # 235.993
mean(sunspot.classic)  # 44.78409
```

#### Evaluate White Noise Terms
```{r}
p = plotts.wge(s2$res)
acf(s2$res, lag.max = 50)
```

```{r}
ljung.wge(s2$res, p = 2) 
```

```{r}
ljung.wge(s2$res, p = 2, K = 48) 
```

* The residuals look reasonably white
* Sample autocorrelation of the residuals stay within the 95% limit lines
* **Ljung-Box did not reject the null of white noise at K=48 (p-value=.21) but at K=48 (p-value=.05) the conclusion of white noise is somewhat questionable**

### AIC Model

#### Model ID
```{r}
aic5.wge(sunspot.classic, p=0:10, q=0:0)
# AIC picks an AR(8)
# FYI BIC selects an AR(2)
```

#### Parameter Estimation
```{r}
s8 = est.ar.wge(sunspot.classic, p=8)
s8$phi  # 1.22872595 -0.47331327 -0.13807811  0.15688938 -0.14030802  0.07050449 -0.12841889 0.20692558
s8$avar # 212.6003
mean(sunspot.classic)  # 44.78409
```

#### Evaluate White Noise Terms
```{r}
p = plotts.wge(s8$res)
acf(s8$res, lag.max = 50)
```

```{r}
ljung.wge(s8$res, p = 8) 
```

```{r}
ljung.wge(s8$res, p = 8, K = 48) 
```

* Again, the residuals look reasonably white
* Sample autocorrelation of the residuals stay within the 95% limit lines
* Ljung-Box did not reject the null of white noise at K=24 or at K=48 with p-values of .33 and .24, respectively 

### Comparing Models

```{r}
# Figure 9.8   4 Realizations from BJR AR(2) model for SS data 
sun2 = est.arma.wge(sunspot.classic,p=2)

s21=gen.arma.wge(n=176, phi=sun2$phi, vara=236,sn=151)+44.78
s22=gen.arma.wge(n=176, phi=sun2$phi, vara=236,sn=161)+44.78
s23=gen.arma.wge(n=176, phi=sun2$phi, vara=236,sn=171)+44.78
s24=gen.arma.wge(n=176, phi=sun2$phi, vara=236,sn=181)+44.78

# Figure 9.9   4 reAalizations from AR(8) model for SS data 
sun8=est.arma.wge(sunspot.classic,p=8)
s81=gen.arma.wge(n=176,phi=sun8$phi,vara=213,sn=938)+44.78
s82=gen.arma.wge(n=176,phi=sun8$phi,vara=213,sn=939)+44.78
s83=gen.arma.wge(n=176,phi=sun8$phi,vara=213,sn=940)+44.78
s84=gen.arma.wge(n=176,phi=sun8$phi,vara=213,sn=941)+44.78

```


#### Realizations

```{r}
numrows <- 2
numcols <- 2
timelab <- 'Time'
valuelab <- ''

fig.width <- 5.2
fig.height <- 3.8
cex.labs <- c(.8,.7,.8)
par(mfrow=c(numrows,numcols),mar=c(3.5,2.5,.8,1))
t<-1:176;

k<-0:10;
ff<-1:251
f<-(ff-1)/500
a.subtitle<-'(a)'
b.subtitle<-'(b)'
c.subtitle<-'(c)'
d.subtitle<-'(d)'

#
plot(t,s21,type='o',xaxt='n',yaxt='n',cex=0.5,pch=16,cex.lab=.75,cex.axis=.75,lwd=.75,xlab='',ylab='',ylim=c(-90,190))
axis(side=1,cex.axis=.9,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.9,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(timelab,valuelab,a.subtitle),line=c(1,1.1,2))
#
plot(t,s22,type='o',xaxt='n',yaxt='n',cex=0.5,pch=16,cex.lab=.75,cex.axis=.75,lwd=.75,xlab='',ylab='',ylim=c(-90,190))
axis(side=1,cex.axis=.9,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.9,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(timelab,valuelab,b.subtitle),line=c(1,1.1,2))
#
plot(t,s23,type='o',xaxt='n',yaxt='n',cex=0.5,pch=16,cex.lab=.75,cex.axis=.75,lwd=.75,xlab='',ylab='',ylim=c(-90,190))
axis(side=1,cex.axis=.9,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.9,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(timelab,valuelab,c.subtitle),line=c(1,1.1,2))
#
plot(t,s24,type='o',xaxt='n',yaxt='n',cex=0.5,pch=16,cex.lab=.75,cex.axis=.75,lwd=.75,xlab='',ylab='',ylim=c(-90,180))
axis(side=1,cex.axis=.9,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.9,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(timelab,valuelab,d.subtitle),line=c(1,1.1,2))
```


```{r}
numrows <- 2
numcols <- 2
timelab <- 'Time'
valuelab <- ''

cex.labs <- c(.8,.7,.8)

par(mfrow=c(numrows,numcols),mar=c(3.5,2.5,.8,1))
t<-1:176;

k<-0:10;
ff<-1:201
f<-(ff-1)/400
a.subtitle<-'(a)'
b.subtitle<-'(b)'
c.subtitle<-'(c)'
d.subtitle<-'(d)'

#
#
plot(t,s81,type='o',xaxt='n',yaxt='n',cex=0.5,pch=16,cex.lab=.75,cex.axis=.75,lwd=.75,xlab='',ylab='',ylim=c(-90,190))
axis(side=1,cex.axis=.9,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.9,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(timelab,valuelab,a.subtitle),line=c(1,1.1,2))
#
plot(t,s82,type='o',xaxt='n',yaxt='n',cex=0.5,pch=16,cex.lab=.75,cex.axis=.75,lwd=.75,xlab='',ylab='',ylim=c(-90,190))
axis(side=1,cex.axis=.9,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.9,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(timelab,valuelab,b.subtitle),line=c(1,1.1,2))
#
plot(t,s83,type='o',xaxt='n',yaxt='n',cex=0.5,pch=16,cex.lab=.75,cex.axis=.75,lwd=.75,xlab='',ylab='',ylim=c(-90,190))
axis(side=1,cex.axis=.9,mgp=c(3,0.15,0),tcl=-.3);9
axis(side=2,las=1,cex.axis=.9,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(timelab,valuelab,c.subtitle),line=c(1,1.1,2))
#
plot(t,s84,type='o',xaxt='n',yaxt='n',cex=0.5,pch=16,cex.lab=.75,cex.axis=.75,lwd=.75,xlab='',ylab='',ylim=c(-90,190))
axis(side=1,cex.axis=.9,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.9,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(timelab,valuelab,d.subtitle),line=c(1,1.1,2))

```

#### Autocorrelations
```{r}
# Figure 9.10   autocorrelation estimates for ss data 

numrows <- 1
numcols <- 2
freqlab <- 'Frequency'
valuelab <- ''
laglab<-'Lag'
a.subtitle<-'(a) Sample autocorrelations for AR(2) realizations'
b.subtitle<-'(b) Sample autocorrelations for AR(8) realizations'

t<-1:100;
k<-0:25
sun=plotts.sample.wge(sunspot.classic)
sun.aut=sun$autplt[1:26]
s21p=plotts.sample.wge(s21)
aut21=s21p$autplt[1:26]
s22p=plotts.sample.wge(s22)
aut22=s22p$autplt[1:26]
s23p=plotts.sample.wge(s23)
aut23=s23p$autplt[1:26]
s24p=plotts.sample.wge(s24)
aut24=s24p$autplt[1:26]


s81p=plotts.sample.wge(s81)
aut81=s81p$autplt[1:26]
s82p=plotts.sample.wge(s82)
aut82=s82p$autplt[1:26]
s83p=plotts.sample.wge(s83)
aut83=s83p$autplt[1:26]
s84p=plotts.sample.wge(s84)
aut84=s84p$autplt[1:26]

k<-0:25;
ff<-1:251
f<-(ff-1)/500

#
#
cex.labs <- c(.67,.7,.67)

par(mfrow=c(numrows,numcols),mar=c(3.8,2.5,1,1))
plot(k,sun.aut,type='o',xaxt='n',yaxt='n',cex=0.65,pch=16,cex.lab=.75,cex.axis=.75,lwd=2,xlab='',ylab='',ylim=c(-1,1))
axis(side=1,cex.axis=.8,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.8,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(laglab,valuelab,a.subtitle),line=c(.8,1.1,1.8))
points(k,aut21,type='o',cex=0.47,pch=1,lwd=.55);
points(k,aut22,type='o',cex=0.47,pch=1,lwd=.55);
points(k,aut23,type='o',cex=0.47,pch=1,lwd=.55);
points(k,aut24,type='o',cex=0.47,pch=1,lwd=.55);


plot(k,sun.aut,type='o',xaxt='n',yaxt='n',cex=0.65,pch=16,cex.lab=.75,cex.axis=.75,lwd=2,xlab='',ylab='',ylim=c(-1,1))
axis(side=1,cex.axis=.8,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.8,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(laglab,valuelab,b.subtitle),line=c(.8,1.1,1.8))
points(k,aut81,type='o',cex=0.47,pch=1,lwd=.55);
points(k,aut82,type='o',cex=0.47,pch=1,lwd=.55);
points(k,aut83,type='o',cex=0.47,pch=1,lwd=.55);
points(k,aut84,type='o',cex=0.47,pch=1,lwd=.55);

```

* AR(8) is better and more consistent with the sunspot data

#### Spectral Densities

```{r}
# Figure 9.11   spectral density estimates for ss data 
numrows <- 1
numcols <- 2
freqlab <- 'Frequency'
valuelab <- ''
a.subtitle<-'(a) Spectral estimates for AR(2) realizations'
b.subtitle<-'(b) Spectral estimates for AR(8) realizations'

cex.labs <- c(.67,.7,.67)

par(mfrow=c(numrows,numcols),mar=c(3.8,2.5,1,1))
t<-1:100;
sunspec=parzen.wge(sunspot.classic,plot=FALSE)
sun.spec=sunspec$pzgram

#
spec.21<-parzen.wge(s21,plot=FALSE)
spec21=spec.21$pzgram
f=spec.21$freq
spec.22<-parzen.wge(s22,plot=FALSE)
spec22=spec.22$pzgram
spec.23<-parzen.wge(s23,plot=FALSE)
spec23=spec.23$pzgram
spec.24<-parzen.wge(s24,plot=FALSE)
spec24=spec.24$pzgram
#
spec.81<-parzen.wge(s81,plot=FALSE)
spec81=spec.81$pzgram
spec.82<-parzen.wge(s82,plot=FALSE)
spec82=spec.82$pzgram
spec.83<-parzen.wge(s83,plot=FALSE)
spec83=spec.83$pzgram
spec.84<-parzen.wge(s84,plot=FALSE)
spec84=spec.84$pzgram
#
#

plot(f,sun.spec,type='l',xaxt='n',yaxt='n',cex=0.4,cex.lab=.75,cex.axis=.75,lwd=2,xlab='',ylab='',ylim=c(-25,10))
axis(side=1,cex.axis=.8,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.8,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(freqlab,valuelab,a.subtitle),line=c(.8,1.1,1.8))
points(f,spec21,type='l',cex=0.1,pch=2,lwd=.55);
points(f,spec22,type='l',cex=0.1,pch=2,lwd=.55);
points(f,spec23,type='l',cex=0.1,pch=2,lwd=.55);
points(f,spec24,type='l',cex=0.1,pch=2,lwd=.55);


plot(f,sun.spec,type='l',xaxt='n',yaxt='n',cex=0.4,cex.lab=.75,cex.axis=.75,lwd=2,xlab='',ylab='',ylim=c(-25,10))
axis(side=1,cex.axis=.8,mgp=c(3,0.15,0),tcl=-.3);
axis(side=2,las=1,cex.axis=.8,mgp=c(3,.4,0),tcl=-.3)
mtext(side=c(1,2,1),cex=cex.labs,text=c(freqlab,valuelab,b.subtitle),line=c(.8,1.1,1.8))
points(f,spec81,type='l',cex=0.1,pch=2,lwd=.55);
points(f,spec82,type='l',cex=0.1,pch=2,lwd=.55);
points(f,spec83,type='l',cex=0.1,pch=2,lwd=.55);
points(f,spec84,type='l',cex=0.1,pch=2,lwd=.55);

```
* Again, AR(8) is better tha the AR(2) as it matches the peak at f = 0 better compared to AR(2)

#### Forecasts

```{r}
f = fore.arma.wge(sunspot.classic, phi = s2$phi, n.ahead = 50, limits = TRUE)
f = fore.arma.wge(sunspot.classic, phi = s8$phi, n.ahead = 50, limits = TRUE)
```

* AR(8) better than AR(2) but both are not realistic

```{r}
f = fore.arma.wge(sunspot.classic, phi = s2$phi, n.ahead = 50, lastn = TRUE, limits = TRUE)
f = fore.arma.wge(sunspot.classic, phi = s8$phi, n.ahead = 50, lastn = TRUE, limits = TRUE)
```

* AR(8) is doing better in the short term, but the further we go, the peaks are predicted at valleys (there is a offset or phse shift) which can be an issue.

### Seasonal Model Fit

```{r}
factor.wge.season(11)
```

```{r}
e = est.ar.wge(sunspot.classic, p = 24, type = 'burg')
```