---
title: "BLS Data"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '6'
  word_document:
    toc: yes
    toc_depth: '6'
always_allow_html: yes
---

* Data from BLS website using code from Jonathan Regenstein
* https://www.linkedin.com/posts/jkregenstein_rstudio-data-quantitativefinance-activity-6631663823331880960-gM4v/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Setup

```{r message=FALSE, warning=FALSE}
library(tswge)
library(tidyverse)
library(tidyquant)
library(jsonlite)
library(httr)
```

# Fetch Data
```{r}
key = readLines(file.choose())
```

```{r}
series = "CES0000000001"  # Seasonally adjusted non farm jobs data
pull = fromJSON(paste0("https://api.bls.gov/publicAPI/v2/timeseries/data/", series,
                       "?registrationkey=", key,
                       "&catalog=true&startyear=2001&endyear=2020")) 

data = pull %>%
  pluck("Results", "series", "data", 1) %>% 
  simplify() %>% 
  select("year", "period", "value") %>% 
  unite("date", c("year", "period"), sep = "-", remove = TRUE) %>% 
  mutate(date = ymd(parse_date_time(date, "%Y%m")),
         jobs = as.numeric(value)) %>% 
  select(-value) %>% 
  arrange(date)

str(data)
head(data)
tail(data)
```

# EDA

```{r}
px = plotts.sample.wge(data$jobs)
```


## Initial Observations
* Note that the realization Y-axis is chopped off. The values are from 130,000 to 150,000.
* ACF plots show extended ACFs pointing to a ARIMA type model.
* Spectral density also shows peak at f = 0 indicating wandering behavior (also seen in the realization).
* This can be removed by applying a High Pass Filter (differencing the data)

## First Difference
```{r}
dif1 = artrans.wge(data$jobs, phi.tr = 1)
px = plotts.sample.wge(dif1)
```

### Observations
* We still see extended ACFs pointing to at least another ARIMA term 
* We will take second difference to to remove the wandering behavior due to ARIMA

## Second Difference
```{r}
dif2 = artrans.wge(dif1, phi.tr = 1)
px = plotts.sample.wge(dif2)
```

### Observations
* Looks like the residual is an MA model with q = 1 (since ACF from lag = 2 onwards is almost 0)
* Let's confirm this by finding the best fit for this residual based on AIC.

## Checking Residuals
```{r}
aic5.wge(dif2)
```

### Observations
* Spot on. Looks like the MA(1) model fits the residual well.
* Full Data seems to fit ARIMA(0,2,1) well.