---
title: "GDP Prediction"
author: "Nikhil Gupta"
date: "`r Sys.time()`"
output:
  github_document: 
    toc: yes
    toc_depth: 6
  word_document:
    toc: yes
    toc_depth: '6'
  html_document:
    toc: yes
    toc_depth: 6
    toc_float: yes
always_allow_html: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Setup
```{r message=FALSE, warning=FALSE}
library(tswge)
library(tswgewrapped)
library(tidyverse)
library(ggplot2)
library(tseries)
library(RColorBrewer)
library(reshape2)
```

```{r}
data = read.csv("../data/economic_indicators_all_ex_3mo_china.csv")
data %>% glimpse()
```

```{r}
data <- data %>% dplyr::select(c(gdp_change, nfjobschg, ipichg, treas10yr, fedintrate, cpichg, inventorieschg, ppichg, popchg, homeownership, corpprofitchg, personincomechg, housingpermitschg))

#data <- data %>% dplyr::select(c(gdp_change, inventorieschg, personincomechg, housingpermitschg))
```

# CCF Analysis

```{r}
lag.max = 12
```

```{r}

var_interest = 'gdp_change'

for (name in colnames(data)){
  # By convention, the X2 variable comes first
  c = ccf(data[name], data[var_interest], lag.max = lag.max)
  index = which(abs(c$acf[,1,1]) == max(abs(c$acf[,1,1])))
  max_ccf = c$lag[,1,1][index] 
  cat(paste("\nFor variable: ", name, " , max cross-correlation is at lag: ", max_ccf, sep = ""))
}


```

* TODO: Add some notes here

* We will use lag.max = 12 in VARselect to give it enough room to pick the best model.


## Backward Variable Selection

We start with the following variables in the model.
AIC is used for selection criteria.


gdp_change
nfjobschg
ipichg 
treas10yr
fedintrate 
cpichg
inventorieschg
ppichg 
popchg
homeownership
corpprofitchg
personincomechg 
housingpermitschg 

### VAR with `trend_type = Trend`

#### Iteration 1

The following variables were selected in iteration one.

personincomechg 
housingpermitschg 
fedintrate 
nfjobschg 
ipichg 

AIC selected a model of lag 12 and lag 12 is used in at least one of the variables.

```{r}
lag.max = 12
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "trend", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data = data, var_interest = var_interest,
                                            mdl_list = models, verbose = 1)

  
mdl_build$summarize_build()
```


#### Iteration 2

The following variables were selected in iteration two

personincomechg 
housingpermitschg 

AIC selected a model of lag 6, but only lag 4 was significant in at least one of the variables.

```{r}
lag.max = 12
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "trend", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data = data %>% dplyr::select(c(gdp_change, personincomechg, housingpermitschg,
                                                                          fedintrate, nfjobschg, ipichg)),
                                          var_interest = var_interest,
                                          mdl_list = models,
                                          verbose = 1)

  
mdl_build$summarize_build()
```


#### Iteration 3

Final iteration is reached with the two variables in the model with at least one significant variable at lag 4.

```{r}
lag.max = 4
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "trend", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data = data %>% dplyr::select(c(gdp_change, personincomechg, housingpermitschg)),
                                          var_interest = var_interest,
                                          mdl_list = models,
                                          verbose = 1)

  
mdl_build$summarize_build()
```


### VAR with `trend_type = Const`

#### Iteration 1

The following variables were selected in iteration one.

nfjobschg 
ipichg 
treas10yr 
fedintrate 
popchg 
homeownership 
corpprofitchg 
personincomechg 
housingpermitschg 

AIC selected a model of lag 12 and lag 12 is used in at least one of the variables.

```{r}
lag.max = 12
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "const", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data = data, var_interest = var_interest,
                                            mdl_list = models, verbose = 1)

  
mdl_build$summarize_build()
```


#### Iteration 2

All variables selected in iteration 1 remain in the model.

AIC selected a model of lag 12, but only lag 11 significant in the model.

```{r}
lag.max = 12
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "const", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data %>% dplyr::select(c(gdp_change, personincomechg, housingpermitschg,
                                                                   fedintrate, nfjobschg, ipichg, treas10yr, popchg,
                                                                   homeownership, corpprofitchg)),
                                           var_interest = var_interest,
                                           mdl_list = models, verbose = 1)

  
mdl_build$summarize_build()
```


#### Iteration 3

All variables selected in iteration 1 remain in the model.

AIC selected a model of lag 11 and at least one variable of lag 11 is significant in the model.

```{r}
lag.max = 11
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "const", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data %>% dplyr::select(c(gdp_change, personincomechg, housingpermitschg,
                                                                   fedintrate, nfjobschg, ipichg, treas10yr, popchg,
                                                                   homeownership, corpprofitchg)),
                                           var_interest = var_interest,
                                           mdl_list = models, verbose = 1)

  
mdl_build$summarize_build()
```



### VAR with `trend_type = Both`

#### Iteration 1

The following variables were selected in iteration one.

nfjobschg 
ipichg  
treas10yr 
fedintrate  
cpichg 
popchg 
homeownership 
personincomechg  
housingpermitschg 

AIC selected a model of lag 12 and lag 12 is used in at least one of the variables.

```{r}
lag.max = 12
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "both", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data = data, var_interest = var_interest,
                                            mdl_list = models, verbose = 1)

  
mdl_build$summarize_build()
```


#### Iteration 2

The following variables were selected in iteration two

nfjobschg 
ipichg  
treas10yr 
fedintrate  
personincomechg  
housingpermitschg 

AIC selected a model of lag 12, but only the variables were only significant to lag 10.

```{r}
lag.max = 12
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "both", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data %>% dplyr::select(c(gdp_change, nfjobschg, ipichg, treas10yr,
                                                                   fedintrate, cpichg, popchg, homeownership,
                                                                   personincomechg, housingpermitschg)),
                                          var_interest = var_interest,
                                          mdl_list = models, verbose = 1)

  
mdl_build$summarize_build()
```


#### Iteration 3

The following variables were selected in iteration three

fedintrate  
personincomechg  
housingpermitschg 

AIC selected a model of lag 6, but only the variables were only significant to lag 2.

```{r}
lag.max = 10
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "both", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data %>% dplyr::select(c(gdp_change, nfjobschg, ipichg, treas10yr,
                                                                   fedintrate, personincomechg, housingpermitschg)),
                                          var_interest = var_interest,
                                          mdl_list = models, verbose = 1)

  
mdl_build$summarize_build()
```

#### Iteration 4

All models remain in the model at itertion 4.

AIC selected a model of lag 2 and all variables are significant at lag of 2.

```{r}
lag.max = 2
  
models = list(
                "VARS AIC Trend"    = list(select = "aic", trend_type = "both", lag.max = lag.max)
                )
var_interest = 'gdp_change'

mdl_build = ModelBuildMultivariateVAR$new(data %>% dplyr::select(c(gdp_change, fedintrate,
                                                                   personincomechg, housingpermitschg)),
                                          var_interest = var_interest,
                                          mdl_list = models, verbose = 1)

  
mdl_build$summarize_build()
```


### Compare Selected Models

```{r}

n.ahead <- 2
batch_size <- 130 ## 12 years to predict the next 2 quarters

# VAR with `trend_type = Trend`

lag.max <- 4 # Increased since AIC is hitting its limit
model.name <- 'VAR_AIC_Trend'
model.variables <- c('gdp_change', 'personincomechg', 'personincomechg')

models = list(
  'VAR_AIC_Trend' = list(type = "VAR", select = "aic", trend_type = "trend", lag.max = lag.max, sliding_ase = TRUE)
)

mdl_compare <- ModelCompareMultivariateVAR$new(data = data %>% dplyr::select(model.variables),
                                              mdl_list = models, var_interest = var_interest,
                                              n.ahead = n.ahead, batch_size = batch_size, verbose = 1)


mdl_comp <- mdl_compare$get_tabular_metrics(ases = FALSE) %>% na.omit()
forecasts <- data.frame(forecast = mdl_comp[which(mdl_comp$Model == model.name), 'f'],
                        model = rep(model.name, length(which(mdl_comp$Model == model.name))),
                        step = which(mdl_comp$Model == model.name) + batch_size)

aes_cmp <- melt(mdl_compare$get_tabular_metrics(ases = T),
                id.vars = c("Model", 'ASE'), measure.vars = c("Time_Test_Start", "Time_Test_End")) %>%
            select(-c(variable))

# VAR with `trend_type = Trend`

lag.max <- 11 # Increased since AIC is hitting its limit
model.name <- 'VAR_AIC_Const'
model.variables <- c('gdp_change', 'nfjobschg', 'ipichg',
                     'treas10yr', 'fedintrate', 'popchg', 
                     'homeownership', 'corpprofitchg', 
                     'personincomechg', 'housingpermitschg')

models = list(
  'VAR_AIC_Const' = list(type = "VAR", select = "aic", trend_type = "const", lag.max = lag.max, sliding_ase = TRUE)
)

mdl_compare <- ModelCompareMultivariateVAR$new(data = data %>% dplyr::select(model.variables),
                                              mdl_list = models, var_interest = var_interest,
                                              n.ahead = n.ahead, batch_size = batch_size, verbose = 1)


mdl_comp <- mdl_compare$get_tabular_metrics(ases = FALSE) %>% na.omit()
forecasts <- rbind(forecasts, data.frame(forecast = mdl_comp[which(mdl_comp$Model == model.name), 'f'],
                                         model = rep(model.name, length(which(mdl_comp$Model == model.name))),
                                         step = which(mdl_comp$Model == model.name) + batch_size))

aes_cmp <- rbind(aes_cmp, melt(mdl_compare$get_tabular_metrics(ases = T),
                               id.vars = c("Model", 'ASE'), measure.vars = c("Time_Test_Start", "Time_Test_End")) %>%
                            select(-c(variable)))

# VAR with `trend_type = Trend`

lag.max <- 2 # Increased since AIC is hitting its limit
model.name <- 'VAR_AIC_Both'
model.variables <- c('gdp_change', 'fedintrate', 'personincomechg', 'personincomechg')

models = list(
  'VAR_AIC_Both' = list(type = "VAR", select = "aic", trend_type = "both", lag.max = lag.max, sliding_ase = TRUE)
)

mdl_compare <- ModelCompareMultivariateVAR$new(data = data %>% dplyr::select(model.variables),
                                              mdl_list = models, var_interest = var_interest,
                                              n.ahead = n.ahead, batch_size = batch_size, verbose = 1)


mdl_comp <- mdl_compare$get_tabular_metrics(ases = FALSE) %>% na.omit()
forecasts <- rbind(forecasts, data.frame(forecast = mdl_comp[which(mdl_comp$Model == model.name), 'f'],
                                         model = rep(model.name, length(which(mdl_comp$Model == model.name))),
                                         step = which(mdl_comp$Model == model.name) + batch_size))

aes_cmp <- rbind(aes_cmp, melt(mdl_compare$get_tabular_metrics(ases = T),
                               id.vars = c("Model", 'ASE'), measure.vars = c("Time_Test_Start", "Time_Test_End")) %>%
                            select(-c(variable)))
```

## Compare Results of Model Selection

### Forecasts vs Realization

```{r}
realization.end <- length(data$gdp_change)
forecast.steps <- length(data$gdp_change[batch_size : realization.end])

forecasts <- rbind(
  forecasts,
  data.frame(
    f = data$gdp_change[batch_size : realization.end],
    model = rep('Realization', forecast.steps),
    step = seq(forecast.steps) + batch_size
    )
  )

forecasts %>%
  ggplot(aes(x = step, y = f, col = model)) +
  geom_line() +
  ggtitle('Model Forecasts with Realization') +
  ylab('Change in GDP') +
  xlab('Time Step')
```

### Forecasts vs Realization (with Reasonable Models)

```{r}
forecasts %>%
  filter(model %in% c('VAR_AIC_Trend', 'VAR_AIC_Both', 'Realization')) %>%
  ggplot(aes(x = step, y = f, col = model)) +
  geom_line() +
  ggtitle('Model Forecasts with Realization') +
  ylab('Change in GDP') +
  xlab('Time Step')
```

### ASE Over Time 

```{r}
aes_cmp %>% ggplot(aes(y = ASE, x = value, col = Model)) +
  geom_line()
```

### ASE Over Time (with Reasonable Models)

```{r}
aes_cmp  %>% 
  filter(Model %in% c('VAR_AIC_Trend', 'VAR_AIC_Both')) %>%
  ggplot(aes(y = ASE, x = value, col = Model)) +
  geom_line()
```

























## Multivariate Compare Object

```{r}

lag.max = 4 # Increased since AIC is hitting its limit

models = list(
  # "VARS AIC Both A" = list(type = "VAR", select = "aic", trend_type = "both", lag.max = lag.max, sliding_ase = FALSE),
  # "VARS BIC Both A" = list(type = "VAR", select = "bic", trend_type = "both", lag.max = lag.max, sliding_ase = FALSE),
  "VARS AIC Both B" = list(type = "VAR", select = "aic", trend_type = "both", lag.max = lag.max, sliding_ase = TRUE),
  #"VARS AIC Trend B" = list(type = "VAR", select = "aic", trend_type = "trend", lag.max = lag.max, sliding_ase = TRUE),
  #"VARS AIC Const B" = list(type = "VAR", select = "aic", trend_type = "const", lag.max = lag.max, sliding_ase = TRUE),
  #"VARS AIC None B" = list(type = "VAR", select = "aic", trend_type = "none", lag.max = lag.max, sliding_ase = TRUE),
  "VARS BIC Both B" = list(type = "VAR", select = "bic", trend_type = "both", lag.max = lag.max, sliding_ase = TRUE)#,
  #"VARS BIC Trend B" = list(type = "VAR", select = "bic", trend_type = "trend", lag.max = lag.max, sliding_ase = TRUE),
  #"VARS BIC Const B" = list(type = "VAR", select = "bic", trend_type = "const", lag.max = lag.max, sliding_ase = TRUE),
  #"VARS BIC None B" = list(type = "VAR", select = "bic", trend_type = "none", lag.max = lag.max, sliding_ase = TRUE)
)
```

```{r}
n.ahead = 2
batch_size = 100 ## 12 years to predict the next 2 quarters
```


```{r}
mdl_compare = ModelCompareMultivariateVAR$new(data = data, mdl_list = models, var_interest = var_interest,
                                              n.ahead = n.ahead, batch_size = batch_size, verbose = 1)
```

### AIC BIC Full Data
```{r}
mdl_compare$get_xIC() 
```

### Tabular Metrics

```{r}
ASEs = mdl_compare$get_tabular_metrics() 

ASEs %>% 
  group_by(Model) %>% 
  summarise(ASE_mean = mean(ASE),
            ASE_median = median(ASE),
            ASE_sd = sd(ASE),
            num_batches = n())

```


```{r}
mdl_compare$get_tabular_metrics(ases = FALSE) %>% 
  na.omit()
```

### Remove unwanted models
```{r}
#mdls_to_remove = c("VARS AIC Both B", "VARS AIC Const B", "VARS AIC None B", "VARS AIC Trend B")
#mdl_compare$remove_models(mdl_names = mdls_to_remove)
```


### Simple Forecasts
```{r fig.width=10}
mdl_compare$plot_simple_forecasts(lastn = TRUE, limits = TRUE)
mdl_compare$plot_simple_forecasts(lastn = FALSE, limits = TRUE)
```


### Batch Forecasts
```{r fig.width=10}
mdl_compare$plot_batch_forecasts(only_sliding = FALSE)
```

### Batch ASEs
```{r fig.width=10}
mdl_compare$plot_batch_ases(only_sliding = FALSE)
```

### ASE Histograms
```{r}
mdl_compare$plot_histogram_ases()  # Same function as univariate
```


### Statistical Compare
```{r}
mdl_compare$statistical_compare() 
```


# Manual Method

```{r}
n = nrow(data)
train_data = data %>% dplyr::filter(row_number() <= (n - n.ahead))
test_data = data %>% dplyr::filter(row_number() > (n - n.ahead))
```


## VARselect
```{r}
library(vars)
# VARselect will select the best K for this VAR model
# Looks at the cross correlation structure (ccf) to figure this out 
vselect = VARselect(data, lag.max = lag.max, type = "both", season = NULL, exogen = NULL)
vselect # Gives AIC values for various K values

k = vselect$selection[["AIC(n)"]]  # BIC = SC(n)
k
```

AIC picks a VAR(10) model

## VAR Model

```{r}
varfit = VAR(train_data, p=k, type="both")
stats::AIC(varfit)
print(summary(varfit)$varresult[[var_interest]])
```

## Predictions

```{r}
preds = stats::predict(varfit, n.ahead=n.ahead)

results = preds$fcst[[var_interest]] %>% 
  dplyr::as_tibble() %>% 
  dplyr::mutate(Time = seq(n-n.ahead+1,n,1)) 

results

```

### ASE

```{r}
ASE_data = data %>% 
  dplyr::mutate(Time = dplyr::row_number()) %>% 
  dplyr::full_join(results, by = "Time") %>% 
  na.omit()

ASE_data
```

```{r}
ASE = mean((ASE_data[[var_interest]] - ASE_data$fcst)^2, na.rm = TRUE)
ASE
```


### Plots

```{r}
results = results %>% 
  dplyr::mutate(!!var_interest := fcst) %>% 
  dplyr::mutate(Model = 'VAR Model') %>% 
  dplyr::select(Model, Time, !!var_interest)

combined_data = data %>% 
  dplyr::mutate(Time = dplyr::row_number()) %>% 
  mutate(Model = 'Realization') %>% 
  dplyr::select(Model, Time, !!var_interest) %>% 
  bind_rows(results) %>% 
  mutate(Model = as.factor(Model))

combined_data
```


```{r fig.width=10}
p = ggplot2::ggplot() + 
  ggplot2::geom_line(data = combined_data, mapping = ggplot2::aes_string(x = "Time", y = var_interest, color = "Model"))

print(p)

plot(seq(1,n,1), data[[var_interest]], type = "b")
points(seq(n-n.ahead+1,n,1), preds$fcst[[var_interest]][,'fcst'], type = "b", pch = 15)
```


```{r fig.width=10}
fanchart(preds, colors = brewer.pal(n = 8, name = "Blues"), names = var_interest) 
```


